---
title: "Statistically Speaking 2026 Week 4"
author: "Josh Garzaniti"
date: "2026-09-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install and load in needed libraries
```{r}
library(nflfastR)
library(nflreadr)
library(tidyverse)
library(stringr)
library(ggplot2)
library(ggdark)
library(caret)
library(xgboost)
library(randomForest)
library(ggrepel)
library(nflplotR)
library(mice)
library(lme4)
library(here)
library(DT)
library(htmlwidgets)
library(gt)
library(purrr)
```
Load in our play by play data
```{r}
pbp = load_pbp(2026)

#Create redzone variable:
pbp = pbp%>%
  mutate(redzone = ifelse(yardline_100 <= 20, 1, 0))

league = pbp
```

## How did the league do?


Create a function to automate pulling stats every week
```{r}

calculate_week_stats = function(week_number) {
  
  # -----------------------------
  # Filter league to selected week
  # -----------------------------
  
  league_week = league%>%
    filter(week == week_number)
  
  
  # -----------------------------
  # Points
  # -----------------------------
  
  team_points = league_week%>%
    group_by(game_id)%>%
    summarise(
      home_team = first(home_team),
      away_team = first(away_team),
      home_points = max(total_home_score, na.rm = TRUE),
      away_points = max(total_away_score, na.rm = TRUE),
      .groups = "drop")
  
  team_points = bind_rows(
    
    team_points %>%
      transmute(
        team = home_team,
        points = home_points),
    
    team_points %>%
      transmute(
        team = away_team,
        points = away_points))
  
  
  # -----------------------------
  # Offense
  # -----------------------------
  
  offense = league_week%>%
    filter(
      !is.na(posteam),
      !play_type %in% c(
        "kickoff",
        "extra_point",
        "field_goal"))%>%
    group_by(posteam)%>%
    summarise(
      
      yards = sum(
        yards_gained,
        na.rm = TRUE),
      
      passing_yards = sum(
        passing_yards,
        na.rm = TRUE),
      
      rushing_yards = sum(
        rushing_yards,
        na.rm = TRUE),
      
      offensive_turnovers = sum(
        interception + fumble_lost,
        na.rm = TRUE),
      
      EPA_generated = sum(
        epa,
        na.rm = TRUE),
      
      EPA_passing = sum(
        epa[play_type == "pass"],
        na.rm = TRUE),
      
      EPA_rushing = sum(
        epa[play_type == "run"],
        na.rm = TRUE),
      
      QB_EPA = sum(
        qb_epa,
        na.rm = TRUE),
      
      redzone_EPA = sum(
        epa[redzone == 1],
        na.rm = TRUE),
      
      redzone_drives = n_distinct(
        drive[redzone == 1]),
      
      redzone_touchdowns = n_distinct(
        drive[
          redzone == 1 &
          touchdown == 1]),
      
      redzone_conversion_rate =
        redzone_touchdowns / redzone_drives,
      
      sacks_allowed = sum(
        sack,
        na.rm = TRUE),
      
      pressures_allowed = sum(
        qb_hit,
        na.rm = TRUE),
      
      incomplete_passes = sum(
        incomplete_pass,
        na.rm = TRUE),
      
      offensive_plays = n(),
      
      pass_plays = sum(
        play_type == "pass",
        na.rm = TRUE),
      
      rush_plays = sum(
        play_type == "run",
        na.rm = TRUE),
      
      sack_rate =
        sacks_allowed / pass_plays,
      
      pressure_allowed_rate =
        pressures_allowed / pass_plays,
      
      incomplete_pass_rate =
        incomplete_passes / pass_plays,
      
      third_downs_faced =
        sum(down == 3, na.rm = TRUE),
      
      third_down_conversion_rate =
        sum(third_down_converted, na.rm = TRUE) /
        third_downs_faced,
      
      EPA_generated_per_play =
        EPA_generated / offensive_plays,
      
      EPA_passing_per_play =
        EPA_passing / pass_plays,
      
      EPA_rushing_per_play =
        EPA_rushing / rush_plays,
      
      QB_EPA_per_play =
        QB_EPA / pass_plays,
      
      off_success_rate =
        sum(success, na.rm = TRUE) /
        offensive_plays,
      
      off_rush_success_rate =
        sum(
          success[play_type == "run"],
          na.rm = TRUE) /
        rush_plays,
      
      off_pass_success_rate =
        sum(
          success[play_type == "pass"],
          na.rm = TRUE) /
        pass_plays,
      
      off_average_air_yards_depth =
        mean(air_yards, na.rm = TRUE),
      
      .groups = "drop")%>%
    rename(team = posteam)
  
  
  # -----------------------------
  # Defense
  # -----------------------------
  
  defense = league_week%>%
    filter(
      !is.na(defteam),
      !play_type %in% c(
        "kickoff",
        "extra_point",
        "field_goal"))%>%
    group_by(defteam)%>%
    summarise(
      
      yards_allowed = sum(
        yards_gained,
        na.rm = TRUE),
      
      passing_yards_allowed = sum(
        passing_yards,
        na.rm = TRUE),
      
      rushing_yards_allowed = sum(
        rushing_yards,
        na.rm = TRUE),
      
      defensive_takeaways = sum(
        interception + fumble_forced,
        na.rm = TRUE),
      
      EPA_given_up = sum(
        epa,
        na.rm = TRUE),
      
      EPA_passing_given = sum(
        epa[play_type == "pass"],
        na.rm = TRUE),
      
      EPA_rushing_given = sum(
        epa[play_type == "run"],
        na.rm = TRUE),
      
      QB_EPA_given = sum(
        qb_epa,
        na.rm = TRUE),
      
      redzone_EPA_given = sum(
        epa[redzone == 1],
        na.rm = TRUE),
      
      redzone_drives_allowed = n_distinct(
        drive[redzone == 1]),
      
      redzone_touchdowns_allowed = n_distinct(
        drive[
          redzone == 1 &
          touchdown == 1]),
      
      redzone_conversion_rate_allowed =
        redzone_touchdowns_allowed /
        redzone_drives_allowed,
      
      sacks_forced = sum(
        sack,
        na.rm = TRUE),
      
      pressures = sum(
        qb_hit,
        na.rm = TRUE),
      
      incomplete_passes_forced = sum(
        incomplete_pass,
        na.rm = TRUE),
      
      defensive_plays = n(),
      
      pass_plays_faced = sum(
        play_type == "pass",
        na.rm = TRUE),
      
      rush_plays_faced = sum(
        play_type == "run",
        na.rm = TRUE),
      
      sack_forced_rate =
        sacks_forced / pass_plays_faced,
      
      pressure_rate =
        pressures / pass_plays_faced,
      
      incomplete_pass_forced_rate =
        incomplete_passes_forced /
        pass_plays_faced,
      
      third_downs_forced =
        sum(down == 3, na.rm = TRUE),
      
      third_down_conversion_rate_allowed =
        sum(third_down_converted, na.rm = TRUE) /
        third_downs_forced,
      
      EPA_given_up_per_play =
        EPA_given_up / defensive_plays,
      
      EPA_passing_given_per_play =
        EPA_passing_given / pass_plays_faced,
      
      EPA_rushing_given_per_play =
        EPA_rushing_given / rush_plays_faced,
      
      QB_EPA_given_per_play =
        QB_EPA_given / pass_plays_faced,
      
      def_success_rate =
        1 - (
          sum(success, na.rm = TRUE) /
          defensive_plays),
      
      def_rush_success_rate =
        1 - (
          sum(
            success[play_type == "run"],
            na.rm = TRUE
          ) /
          rush_plays_faced),
      
      def_pass_success_rate =
        1 - (
          sum(
            success[play_type == "pass"],
            na.rm = TRUE) /
          pass_plays_faced),
      
      def_average_air_yards_depth =
        mean(air_yards, na.rm = TRUE),
      
      .groups = "drop")%>%
    rename(team = defteam)
  
  
  # -----------------------------
  # Join offense + defense + points
  # -----------------------------
  
  weekly_summary = offense%>%
    left_join(
      defense,
      by = "team")%>%
    left_join(
      team_points,
      by = "team")%>%
    mutate(
      
      Net_EPA =
        EPA_generated - EPA_given_up,
      
      Net_EPA_per_play =
        EPA_generated_per_play -
        EPA_given_up_per_play,
      
      turnover_differential =
        defensive_takeaways -
        offensive_turnovers,
      
      week = week_number)
  
  
  return(weekly_summary)
}

week_4_stats = calculate_week_stats(4)
```


Map out stats to their respective weeks
```{r}

all_weekly_stats = map(
  sort(unique(league$week)),
  calculate_week_stats)%>%
  list_rbind()%>%
  arrange(week, team)
```

Create the Team Ranks for each stat
```{r}
all_weekly_stats = all_weekly_stats%>%
  group_by(week)%>%
  mutate(
    
    points_rank =
      min_rank(desc(points)),
    
    offensive_plays_rank =
      min_rank(desc(offensive_plays)),
    
    defensive_plays_rank =
      min_rank(defensive_plays),
    
    yards_rank =
      min_rank(desc(yards)),
    
    passing_yards_rank =
      min_rank(desc(passing_yards)),
    
    rushing_yards_rank =
      min_rank(desc(rushing_yards)),
    
    offensive_turnovers_rank =
      min_rank(offensive_turnovers),
    
    EPA_generated_rank =
      min_rank(desc(EPA_generated)),
    
    EPA_passing_rank =
      min_rank(desc(EPA_passing)),
    
    EPA_rushing_rank =
      min_rank(desc(EPA_rushing)),
    
    QB_EPA_rank =
      min_rank(desc(QB_EPA)),
    
    redzone_EPA_rank =
      min_rank(desc(redzone_EPA)),
    
    redzone_conversion_rate_rank =
      min_rank(desc(redzone_conversion_rate)),
    
    sacks_allowed_rank =
      min_rank(sacks_allowed),
    
    sack_rate_rank =
      min_rank(sack_rate),
    
    pressures_allowed_rank =
      min_rank(pressures_allowed),
    
    pressure_allowed_rate_rank =
      min_rank(pressure_allowed_rate),
    
    incomplete_pass_rate_rank =
      min_rank(incomplete_pass_rate),
    
    third_down_conversion_rate_rank =
      min_rank(desc(third_down_conversion_rate)),
    
    EPA_generated_per_play_rank =
      min_rank(desc(EPA_generated_per_play)),
    
    EPA_passing_per_play_rank =
      min_rank(desc(EPA_passing_per_play)),
    
    EPA_rushing_per_play_rank =
      min_rank(desc(EPA_rushing_per_play)),
    
    QB_EPA_per_play_rank =
      min_rank(desc(QB_EPA_per_play)),
    
    off_success_rate_rank =
      min_rank(desc(off_success_rate)),
    
    off_rush_success_rate_rank =
      min_rank(desc(off_rush_success_rate)),
    
    off_pass_success_rate_rank =
      min_rank(desc(off_pass_success_rate)),
    
    off_average_air_yards_depth_rank =
      min_rank(desc(off_average_air_yards_depth)),
    
    yards_allowed_rank =
      min_rank(yards_allowed),
    
    passing_yards_allowed_rank =
      min_rank(passing_yards_allowed),
    
    rushing_yards_allowed_rank =
      min_rank(rushing_yards_allowed),
    
    defensive_takeaways_rank =
      min_rank(desc(defensive_takeaways)),
    
    EPA_given_up_rank =
      min_rank(EPA_given_up),
    
    EPA_passing_given_rank =
      min_rank(EPA_passing_given),
    
    EPA_rushing_given_rank =
      min_rank(EPA_rushing_given),
    
    QB_EPA_given_rank =
      min_rank(QB_EPA_given),
    
    redzone_EPA_given_rank =
      min_rank(redzone_EPA_given),
    
    redzone_conversion_rate_allowed_rank =
      min_rank(redzone_conversion_rate_allowed),
    
    sacks_forced_rank =
      min_rank(desc(sacks_forced)),
    
    sack_forced_rate_rank =
      min_rank(desc(sack_forced_rate)),
    
    pressures_rank =
      min_rank(desc(pressures)),
    
    pressure_rate_rank =
      min_rank(desc(pressure_rate)),
    
    incomplete_pass_forced_rate_rank =
      min_rank(desc(incomplete_pass_forced_rate)),
    
    third_down_conversion_rate_allowed_rank =
      min_rank(desc(third_down_conversion_rate_allowed)),
    
    EPA_given_up_per_play_rank =
      min_rank(EPA_given_up_per_play),
    
    EPA_passing_given_per_play_rank =
      min_rank(EPA_passing_given_per_play),
    
    EPA_rushing_given_per_play_rank =
      min_rank(EPA_rushing_given_per_play),
    
    QB_EPA_given_per_play_rank =
      min_rank(QB_EPA_given_per_play),
    
    def_success_rate_rank =
      min_rank(desc(def_success_rate)),
    
    def_rush_success_rate_rank =
      min_rank(desc(def_rush_success_rate)),
    
    def_pass_success_rate_rank =
      min_rank(desc(def_pass_success_rate)),
    
    def_average_air_yards_depth_rank =
      min_rank(def_average_air_yards_depth),
    
    Net_EPA_rank =
      min_rank(desc(Net_EPA)),
    
    Net_EPA_per_play_rank =
      min_rank(desc(Net_EPA_per_play)),
    
    turnover_differential_rank =
      min_rank(desc(turnover_differential)))%>%
  ungroup()
```


```{r}
league_wk_4_summary = all_weekly_stats%>%
  filter(week == 4)
```

##Segmenting our Stats into offense and defense

Offense
```{r}
offensive_stats = c(
  "points",
  "offensive_plays",
  "yards",
  "passing_yards",
  "rushing_yards",
  "turnover_differential",
  "offensive_turnovers",
  "Net_EPA",
  "Net_EPA_per_play",
  "EPA_generated",
  "EPA_passing",
  "EPA_rushing",
  "QB_EPA",
  "redzone_EPA",
  "redzone_conversion_rate",
  "sacks_allowed",
  "sack_rate",
  "pressures_allowed",
  "pressure_allowed_rate",
  "incomplete_pass_rate",
  "third_down_conversion_rate",
  "EPA_generated_per_play",
  "EPA_passing_per_play",
  "EPA_rushing_per_play",
  "QB_EPA_per_play",
  "off_success_rate",
  "off_rush_success_rate",
  "off_pass_success_rate",
  "off_average_air_yards_depth")

offensive_lower_is_better = c(
  "offensive_turnovers",
  "sacks_allowed",
  "sack_rate",
  "pressures_allowed",
  "pressure_allowed_rate",
  "incomplete_pass_rate")

heatmap_data = league_wk_4_summary%>%
  select(team, all_of(offensive_stats))%>%
  pivot_longer(
    cols = -team,
    names_to = "stat",
    values_to = "value")%>%
  group_by(stat)%>%
  mutate(
    league_avg = mean(value, na.rm = TRUE),
    vs_league_avg = value - league_avg)%>%
  ungroup()
```

Defense
```{r}
defensive_stats = c(
  "defensive_plays",
  "defensive_takeaways",
  "yards_allowed",
  "passing_yards_allowed",
  "rushing_yards_allowed",
  "EPA_given_up",
  "EPA_passing_given",
  "EPA_rushing_given",
  "QB_EPA_given",
  "redzone_EPA_given",
  "redzone_conversion_rate_allowed",
  "sacks_forced",
  "sack_forced_rate",
  "pressures",
  "pressure_rate",
  "incomplete_pass_forced_rate",
  "third_down_conversion_rate_allowed",
  "EPA_given_up_per_play",
  "EPA_passing_given_per_play",
  "EPA_rushing_given_per_play",
  "QB_EPA_given_per_play",
  "def_success_rate",
  "def_rush_success_rate",
  "def_pass_success_rate",
  "def_average_air_yards_depth")

defensive_lower_is_better = c(
  "defensive_plays",
  "yards_allowed",
  "passing_yards_allowed",
  "rushing_yards_allowed",
  "EPA_given_up",
  "EPA_passing_given",
  "EPA_rushing_given",
  "QB_EPA_given",
  "redzone_EPA_given",
  "redzone_conversion_rate_allowed",
  "third_down_conversion_rate_allowed",
  "EPA_given_up_per_play",
  "EPA_passing_given_per_play",
  "EPA_rushing_given_per_play",
  "QB_EPA_given_per_play",
  "def_average_air_yards_depth")

all_stats = c(
 offensive_stats,
  defensive_stats)

heatmap_stats = c(
  offensive_stats,
  defensive_stats)

heatmap_data = league_wk_4_summary%>%
  select(team, all_of(heatmap_stats))%>%
  pivot_longer(
    cols = -team,
    names_to = "stat",
    values_to = "value")%>%
  group_by(stat) %>%
  mutate(
    league_avg = mean(value, na.rm = TRUE),
    vs_league_avg = value - league_avg)%>%
  ungroup() %>%
mutate(
  performance_vs_avg = case_when(
    stat %in% defensive_lower_is_better ~ -vs_league_avg,
    stat %in% offensive_lower_is_better ~ -vs_league_avg,
    TRUE ~ vs_league_avg))%>%
  group_by(stat)%>%
  mutate(
    z_score = (
      performance_vs_avg -
        mean(performance_vs_avg, na.rm = TRUE)) /
      sd(performance_vs_avg, na.rm = TRUE))%>%
  ungroup()

```


Offensive Heatmap
```{r}
offensive_heatmap_data = heatmap_data%>%
  filter(stat %in% offensive_stats)

off_heatmap_wk4 = ggplot(
  offensive_heatmap_data,
  aes(
    x = team,
    y = stat,
    fill = z_score))+
  
  geom_tile(color = "grey20")+
  
  scale_fill_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = 0)+
  
  labs(
    title = "NFL Week 4 Offensive Performance vs. League Average",
    subtitle = "Positive values indicate above-average performance",
    x = "Team",
    y = "Statistic",
    fill = "Vs. League Avg.")+
  
  theme_minimal()+
  theme(
    plot.background = element_rect(fill = "black", color = NA),
    panel.background = element_rect(fill = "black", color = NA),
    panel.grid = element_blank(),
    
    text = element_text(color = "white"),
    axis.text = element_text(color = "white"),
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      color = "white"),
    axis.title = element_text(color = "white"),
    
    plot.title = element_text(
      color = "white",
      face = "bold"),
    plot.subtitle = element_text(color = "grey80"),
    
    legend.background = element_rect(fill = "black", color = NA),
    legend.text = element_text(color = "white"),
    legend.title = element_text(color = "white"))

off_heatmap_wk4
```

Saving Week 4 Heatmap
```{r}
ggsave(
  filename = "C:/Personal Projects/Statistically Speaking/NFL_Week_4_Offensive_Heatmap.png",
  plot = off_heatmap_wk4,
  width = 14,
  height = 8,
  dpi = 300,
  bg = "black")
```


Defensive Heatmap
```{r}

defensive_heatmap_data = heatmap_data%>%
  filter(stat %in% defensive_stats)

def_heatmap_wk4 = ggplot(
  defensive_heatmap_data,
  aes(
    x = team,
    y = stat,
    fill = z_score))+
  
  geom_tile(color = "grey20")+
  
  scale_fill_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = 0)+
  
  labs(
    title = "NFL Week 4 Defensive Performance vs. League Average",
    subtitle = "Positive values indicate above-average performance",
    x = "Team",
    y = "Statistic",
    fill = "Vs. League Avg.")+
  
  theme_minimal()+
  theme(
    plot.background = element_rect(fill = "black", color = NA),
    panel.background = element_rect(fill = "black", color = NA),
    panel.grid = element_blank(),
    
    text = element_text(color = "white"),
    axis.text = element_text(color = "white"),
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      color = "white"),
    axis.title = element_text(color = "white"),
    
    plot.title = element_text(
      color = "white",
      face = "bold"),
    plot.subtitle = element_text(color = "grey80"),
    
    legend.background = element_rect(fill = "black", color = NA),
    legend.text = element_text(color = "white"),
    legend.title = element_text(color = "white"))

def_heatmap_wk4
```

Saving Week 4 Heatmap
```{r}
ggsave(
  filename = "C:/Personal Projects/Statistically Speaking/NFL_Week_4_Defensive_Heatmap.png",
  plot = def_heatmap_wk4,
  width = 14,
  height = 8,
  dpi = 300,
  bg = "black")
```


Offensive Charts:
```{r}

off_rush_avg = mean(league_wk_4_summary$EPA_rushing, na.rm = TRUE)
off_pass_avg = mean(league_wk_4_summary$EPA_passing, na.rm = TRUE)

off_epa_wk_4 = ggplot(
  league_wk_4_summary,
  aes(
    x = EPA_rushing,
    y = EPA_passing))+
  
  nflplotR::geom_nfl_logos(
    aes(team_abbr = team),
    width = 0.075)+
  
  geom_vline(
    xintercept = off_rush_avg,
    linetype = "dashed",
    color = "grey50")+
  
  geom_hline(
    yintercept = off_pass_avg,
    linetype = "dashed",
    color = "grey50")+

  
  labs(
    title = "Team Offensive Efficiency",
    subtitle = "2026 Week 4 | Higher EPA is Better",
    x = "Rushing EPA",
    y = "Passing EPA")+
  
  theme_minimal()+
  theme(
    plot.background = element_rect(fill = "black", color = NA),
    panel.background = element_rect(fill = "black", color = NA),
    panel.grid = element_blank(),
    
    text = element_text(color = "white"),
    axis.text = element_text(color = "white"),
    axis.title = element_text(color = "white"),
    
    plot.title = element_text(
      color = "white",
      face = "bold"),
    plot.subtitle = element_text(color = "grey80"))

off_epa_wk_4
```

Saving chart
```{r}
ggsave(
  filename = "C:/Personal Projects/Statistically Speaking/NFL_Week_4_Offensive EPA Chart.png",
  plot = off_epa_wk_4,
  width = 14,
  height = 8,
  dpi = 300,
  bg = "black")
```


Success Rate
```{r}

off_rush_suc_avg = mean(league_wk_4_summary$off_rush_success_rate, na.rm = TRUE)
off_pass_suc_avg = mean(league_wk_4_summary$off_pass_success_rate, na.rm = TRUE)

off_suc_wk_4 = ggplot(
  league_wk_4_summary,
  aes(
    x = off_rush_success_rate,
    y = off_pass_success_rate))+
  
  nflplotR::geom_nfl_logos(
    aes(team_abbr = team),
    width = 0.075)+
  
  geom_vline(
    xintercept = off_rush_suc_avg,
    linetype = "dashed",
    color = "grey50")+
  
  geom_hline(
    yintercept = off_pass_suc_avg,
    linetype = "dashed",
    color = "grey50")+

  
  labs(
    title = "Team Offensive Success",
    subtitle = "2026 Week 4 | Higher Success Rate is Better",
    x = "Rushing Success %",
    y = "Passing Success %")+
  
  theme_minimal()+
  theme(
    plot.background = element_rect(fill = "black", color = NA),
    panel.background = element_rect(fill = "black", color = NA),
    panel.grid = element_blank(),
    
    text = element_text(color = "white"),
    axis.text = element_text(color = "white"),
    axis.title = element_text(color = "white"),
    
    plot.title = element_text(
      color = "white",
      face = "bold"),
    plot.subtitle = element_text(color = "grey80"))

off_suc_wk_4
```

Saving chart
```{r}
ggsave(
  filename = "C:/Personal Projects/Statistically Speaking/NFL_Week_4_Offensive Success Rate Chart.png",
  plot = off_suc_wk_4,
  width = 14,
  height = 8,
  dpi = 300,
  bg = "black")
```


Defensive Charts:
```{r}

def_rush_avg = mean(league_wk_4_summary$EPA_rushing_given, na.rm = TRUE)
def_pass_avg = mean(league_wk_4_summary$EPA_passing_given, na.rm = TRUE)

def_epa_wk_4 = ggplot(
  league_wk_4_summary,
  aes(
    x = EPA_rushing_given,
    y = EPA_passing_given))+
  
  nflplotR::geom_nfl_logos(
    aes(team_abbr = team),
    width = 0.075)+
  
  geom_vline(
    xintercept = def_rush_avg,
    linetype = "dashed",
    color = "grey50")+
  
  geom_hline(
    yintercept = def_pass_avg,
    linetype = "dashed",
    color = "grey50")+
  
  scale_x_reverse()+
  scale_y_reverse()+
  
  labs(
    title = "Team Defensive Efficiency",
    subtitle = "2026 Week 4 | Lower EPA Allowed is Better",
    x = "Rushing EPA Allowed",
    y = "Passing EPA Allowed")+
  
  theme_minimal()+
  theme(
    plot.background = element_rect(fill = "black", color = NA),
    panel.background = element_rect(fill = "black", color = NA),
    panel.grid = element_blank(),
    
    text = element_text(color = "white"),
    axis.text = element_text(color = "white"),
    axis.title = element_text(color = "white"),
    
    plot.title = element_text(
      color = "white",
      face = "bold"),
    plot.subtitle = element_text(color = "grey80"))

def_epa_wk_4
```

Saving chart
```{r}
ggsave(
  filename = "C:/Personal Projects/Statistically Speaking/NFL_Week_4_Defensive EPA Chart.png",
  plot = def_epa_wk_4,
  width = 14,
  height = 8,
  dpi = 300,
  bg = "black")
```



Defensive Success Rate
```{r}

def_rush_suc_avg = mean(league_wk_4_summary$def_rush_success_rate, na.rm = TRUE)
def_pass_suc_avg = mean(league_wk_4_summary$def_pass_success_rate, na.rm = TRUE)

def_suc_wk_4 = ggplot(
  league_wk_4_summary,
  aes(
    x = def_rush_success_rate,
    y = def_pass_success_rate))+
  
  nflplotR::geom_nfl_logos(
    aes(team_abbr = team),
    width = 0.075)+
  
  geom_vline(
    xintercept = def_rush_suc_avg,
    linetype = "dashed",
    color = "grey50")+
  
  geom_hline(
    yintercept = def_pass_suc_avg,
    linetype = "dashed",
    color = "grey50")+

  
  labs(
    title = "Team Defensive Success",
    subtitle = "2026 Week 4 | Higher Success Rate is Better",
    x = "Rushing Success %",
    y = "Passing Success %")+
  
  theme_minimal()+
  theme(
    plot.background = element_rect(fill = "black", color = NA),
    panel.background = element_rect(fill = "black", color = NA),
    panel.grid = element_blank(),
    
    text = element_text(color = "white"),
    axis.text = element_text(color = "white"),
    axis.title = element_text(color = "white"),
    
    plot.title = element_text(
      color = "white",
      face = "bold"),
    plot.subtitle = element_text(color = "grey80"))

def_suc_wk_4
```

Saving chart
```{r}
ggsave(
  filename = "C:/Personal Projects/Statistically Speaking/NFL_Week_4_Defensive Success Rate Chart.png",
  plot = def_suc_wk_4,
  width = 14,
  height = 8,
  dpi = 300,
  bg = "black")
```


#Offensive Ranking
```{r}
offensive_ranking = offensive_heatmap_data%>%
  group_by(stat)%>%
  mutate(
    rank = ifelse(
      stat %in% offensive_lower_is_better,
      min_rank(value),
      min_rank(desc(value))))%>%
  ungroup()%>%
  select(team, stat, rank)%>%
  pivot_wider(
    names_from = stat,
    values_from = rank,
    names_glue = "{stat}_rank")%>%
  mutate(
    average_rank = rowMeans(
      select(., ends_with("_rank")),
      na.rm = TRUE),
    final_rank = min_rank(average_rank))%>%
  arrange(final_rank)
```

Offensive Ranking Table
```{r}
offensive_ranking_table = DT::datatable(
  offensive_ranking,
  options = list(
    scrollX = TRUE,
    pageLength = 10))

offensive_ranking_table
```

Saving table
```{r}
saveWidget(
  offensive_ranking_table,
  "C:/Personal Projects/Statistically Speaking/offensive_ranking_table_wk4.html",
  selfcontained = TRUE)
```


#Defensive Ranking
```{r}
defensive_ranking = defensive_heatmap_data%>%
  group_by(stat)%>%
  mutate(
    rank = ifelse(
      stat %in% defensive_lower_is_better,
      min_rank(value),
      min_rank(desc(value))))%>%
  ungroup()%>%
  select(team, stat, rank)%>%
  pivot_wider(
    names_from = stat,
    values_from = rank,
    names_glue = "{stat}_rank")%>%
  mutate(
    average_rank = rowMeans(
      select(., ends_with("_rank")),
      na.rm = TRUE),
    final_rank = min_rank(average_rank))%>%
  arrange(final_rank)
```

Defensive Ranking Table
```{r}
defensive_ranking_table = DT::datatable(
  defensive_ranking,
  options = list(
    scrollX = TRUE,
    pageLength = 10))

defensive_ranking_table
```

Saving table
```{r}
saveWidget(
  defensive_ranking_table,
  "C:/Personal Projects/Statistically Speaking/defensive_ranking_table_wk4.html",
  selfcontained = TRUE)
```

#Comparing Weekly Performance

Create a function to make comparison between chargers stats week by week
```{r}
compare_weeks = function(
    data,
    team_name = "LAC",
    earlier_week,
    later_week) {
  
  data %>%
    filter(
      team == team_name,
      week %in% c(earlier_week, later_week))%>%
    select(
      team,
      week,
      all_of(all_stats))%>%
    pivot_longer(
      cols = all_of(all_stats),
      names_to = "Stat",
      values_to = "value")%>%
    pivot_wider(
      names_from = week,
      values_from = value,
      names_prefix = "Week_")%>%
    mutate(
      Difference =
        .data[[paste0("Week_", later_week)]] -
        .data[[paste0("Week_", earlier_week)]],
      
      Direction = case_when(
        Difference > 0 ~ "↑",
        Difference < 0 ~ "↓",
        TRUE ~ "→"))%>%
    select(
      Stat,
      all_of(paste0("Week_", earlier_week)),
      all_of(paste0("Week_", later_week)),
      Difference,
      Direction)
}
```

Chargers Week 3 -> week 4 Comparison
```{r}
Chargers_wk3_wk4_table = compare_weeks(
  all_weekly_stats,
  team_name = "LAC",
  earlier_week = 3,
  later_week = 4)
```

Table Visualization of the Comparison
```{r}
Chargers_wk3_wk4_gt =
  Chargers_wk3_wk4_table%>%
  gt()%>%
  tab_header(
    title = "Chargers: Week 3 vs. Week 4",
    subtitle = "Positive differences indicate improved performance")%>%
  cols_label(
    Stat = "Statistic",
    Week_3 = "Week 3",
    Week_4 = "Week 4",
    Difference = "Difference",
    Direction = "Direction")%>%
  fmt_number(
    columns = c(
      Week_3,
      Week_4,
      Difference),
    decimals = 2)%>%
  cols_align(
    align = "center",
    columns = c(
      Week_3,
      Week_4,
      Difference,
      Direction))%>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(everything()))%>%
  tab_style(
    style = cell_text(
      size = px(20),
      weight = "bold"),
    locations = cells_body(
      columns = Direction))%>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = Stat))%>%
  opt_row_striping()%>%
  opt_table_font(
    font = list(
      google_font("Roboto"),
      default_fonts()))%>%
  tab_options(
    table.font.size = px(14),
    heading.title.font.size = px(22),
    heading.subtitle.font.size = px(14))

Chargers_wk3_wk4_gt
```


Save the Comparison Table
```{r}
gtsave(
  Chargers_wk3_wk4_gt,
  "C:/Personal Projects/Statistically Speaking/Chargers_wk3_wk4.png")
```


Chargers vs League Seasonal Comparison

```{r}
current_week = max(league$week)

league_through_current_week =
  all_weekly_stats%>%
  filter(week <= current_week)

league_avg = all_weekly_stats%>%
  group_by(week)%>%
  summarise(
    across(
      all_of(all_stats),
      ~ mean(.x, na.rm = TRUE)),
    .groups = "drop")

chargers_weekly = all_weekly_stats%>%
  filter(team == "LAC")%>%
  select(
    week,
    all_of(all_stats))
```

Chargers Season Avg:
```{r}
chargers_season_avg = all_weekly_stats%>%
  filter(
    team == "LAC",
    week <= current_week)%>%
  summarise(
    across(
      all_of(all_stats),
      ~ mean(.x, na.rm = TRUE)))%>%
  pivot_longer(
    cols = everything(),
    names_to = "stat",
    values_to = "Chargers_Season_Avg")
```

League Season Avg:
```{r}
league_season_avg = all_weekly_stats%>%
  filter(week <= current_week)%>%
  summarise(
    across(
      all_of(all_stats),
      ~ mean(.x, na.rm = TRUE)))%>%
  pivot_longer(
    cols = everything(),
    names_to = "stat",
    values_to = "League_Season_Avg")
```

Difference:
```{r}
league_vs_chargers_season = all_weekly_stats%>%
  filter(week <= current_week)%>%
  select(
    team,
    week,
    all_of(all_stats))%>%
  pivot_longer(
    cols = all_of(all_stats),
    names_to = "stat",
    values_to = "value")%>%
  group_by(stat, week)%>%
  summarise(
    league_avg = mean(
      value,
      na.rm = TRUE),
    .groups = "drop")%>%
  pivot_wider(
    names_from = week,
    values_from = league_avg,
    names_glue = "League_Wk{week}")
```

Isolate Chargers
```{r}
chargers_comparison = all_weekly_stats%>%
  filter(
    team == "LAC",
    week <= current_week)%>%
  select(
    week,
    all_of(all_stats))%>%
  pivot_longer(
    cols = all_of(all_stats),
    names_to = "stat",
    values_to = "value")%>%
  pivot_wider(
    names_from = week,
    values_from = value,
    names_glue = "Chargers_Wk{week}")
```


Create the DF
```{r}
league_vs_chargers_season =
  league_vs_chargers_season%>%
  left_join(
    chargers_comparison,
    by = "stat")
```








