---
title: "Statistically Speaking 2026 Week 2"
author: "Josh Garzaniti"
date: "2026-09-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install and load in needed libraries
```{r}
library(nflfastR)
library(nflreadr)
library(tidyverse)
library(stringr)
library(ggplot2)
library(ggdark)
library(caret)
library(xgboost)
library(randomForest)
library(ggrepel)
library(nflplotR)
library(mice)
library(lme4)
library(here)
library(DT)
library(htmlwidgets)
library(gt)
```
Load in our play by play data
```{r}
pbp = load_pbp(2026)
```

First I want to separate into 2 dataframes: Chargers and the rest of the league

```{r}
chargers = pbp%>%
  filter(home_team == "LAC" | away_team == "LAC")

league = pbp

chargers_offense = chargers%>%
  filter(posteam == "LAC" & !play_type %in% c("kickoff", "extra_point", "field_goal"))

chargers_defense = chargers%>%
  filter(defteam == "LAC" & !play_type %in% c("kickoff", "extra_point", "field_goal"))

chargers_st = chargers%>%
  filter(play_type %in% c("kickoff", "extra_point", "field_goal"))
```


Week 2 vs Las Vegas Raiders

```{r}
chargers_wk_2 = chargers%>%
  filter(posteam == "LAC" 
         & week == 2)

chargers_offense_wk_2 = chargers%>%
  filter(posteam == "LAC" 
         & week == 2
         & !play_type %in% c("kickoff", "extra_point", "field_goal"))

chargers_defense_wk_2 = chargers%>%
  filter(defteam == "LAC"
         & week == 2
         & !play_type %in% c("kickoff", "extra_point", "field_goal"))

chargers_st_wk_2 = chargers%>%
  filter(play_type %in% c("kickoff", "extra_point", "field_goal")
         & week == 2)
```


Summary:

```{r}
total_wk2_points = max(chargers$total_home_score, na.rm = TRUE)

total_wk2_points
  
total_wk2_points_allowed = max(chargers$total_away_score, na.rm = TRUE)

total_wk2_points_allowed

total_wk2_yards = sum(chargers_offense_wk_2$yards_gained, na.rm = TRUE)

total_wk2_yards

total_wk2_yards_allowed = sum(chargers_defense_wk_2$yards_gained, na.rm = TRUE)

total_wk2_yards_allowed

total_wk2_turnover_dif = (sum(chargers_defense_wk_2$interception, chargers_defense_wk_2$fumble_forced, na.rm = TRUE) - sum(chargers_offense_wk_2$interception, chargers_offense_wk_2$fumble_lost, na.rm = TRUE))

total_wk2_turnover_dif

total_wk2_EPA_generated = sum(chargers_offense_wk_2$epa, na.rm = TRUE)

total_wk2_EPA_generated

total_wk2_EPA_given_up = sum(chargers_defense$epa, na.rm = TRUE)

total_wk2_EPA_given_up

total_wk2_Net_EPA = total_wk2_EPA_generated - total_wk2_EPA_given_up

total_wk2_Net_EPA

total_wk2_EPA_generated_per_play = total_wk2_EPA_generated/nrow(chargers_offense_wk_2)

total_wk2_EPA_generated_per_play

total_wk2_EPA_given_up_per_play = total_wk2_EPA_given_up/nrow(chargers_defense_wk_2)

total_wk2_EPA_given_up_per_play

total_wk2_Net_EPA_per_play = total_wk2_EPA_generated_per_play - total_wk2_EPA_given_up_per_play

total_wk2_Net_EPA_per_play

```

Chargers 14
Raiders 26
Chargers yards: 310
Raiders yards: 305
Turnover differential: -1
EPA generated: -19.12942
EPA allowed: -5.740576
Net EPA: -13.38885
EPA generated per play: -0.2484341
EPA allowed per play:-0.07863802
Net EPA per play: -0.1697961


## How did the league do?

```{r}
league_wk_2 = league%>%
  filter(week == 2)
```

Generate Stats for each team:

Scoring Week 2
```{r}
team_points_wk_2 = league_wk_2%>%
  group_by(game_id)%>%
  summarise(
    home_team = first(home_team),
    away_team = first(away_team),
    home_points = max(total_home_score, na.rm = TRUE),
    away_points = max(total_away_score, na.rm = TRUE),
    .groups = "drop")

team_points_wk_2 = bind_rows(
  team_points_wk_2%>%
    transmute(
      team = home_team,
      points = home_points),
  
  team_points_wk_2%>%
    transmute(
      team = away_team,
      points = away_points))
```

Offensive Stats
```{r}
offense_wk_2 = league_wk_2%>%
  filter(
    !is.na(posteam),
    !play_type %in% c(
      "kickoff",
      "extra_point",
      "field_goal"))%>%
  group_by(posteam)%>%
  summarise(
    yards = sum(yards_gained, na.rm = TRUE),
    
    passing_yards = sum(passing_yards, na.rm = TRUE),
    
    rushing_yards = sum(rushing_yards, na.rm = TRUE),
    
    offensive_turnovers = sum(
      interception + fumble_lost,
      na.rm = TRUE),
    

    EPA_generated = sum(epa, na.rm = TRUE),
    

    EPA_passing = sum(
      epa[play_type == "pass"],
      na.rm = TRUE),
    
 
    EPA_rushing = sum(
      epa[play_type == "run"],
      na.rm = TRUE),
    
 
    QB_EPA = sum(qb_epa, na.rm = TRUE),
    
 
    offensive_plays = n(),
    
    pass_plays = sum(play_type == "pass", na.rm = TRUE),
    
    rush_plays = sum(play_type == "run", na.rm = TRUE),
    
 
    EPA_generated_per_play =
      EPA_generated / offensive_plays,
    
    EPA_passing_per_play =
      EPA_passing / pass_plays,
    
    EPA_rushing_per_play =
      EPA_rushing / rush_plays,
    
    QB_EPA_per_play =
      QB_EPA / pass_plays,
    
    .groups = "drop")%>%
  rename(team = posteam)
```

Defensive Stats
```{r}
defense_wk_2 = league_wk_2%>%
  filter(
    !is.na(defteam),
    !play_type %in% c(
      "kickoff",
      "extra_point",
      "field_goal"))%>%
  group_by(defteam)%>%
  summarise(
    yards_allowed = sum(yards_gained, na.rm = TRUE),
    
    passing_yards_allowed = sum(passing_yards, na.rm = TRUE),
    
    rushing_yards_allowed = sum(rushing_yards, na.rm = TRUE),
    
    defensive_turnovers = sum(
  interception + fumble_forced,
  na.rm = TRUE),
  
  EPA_given_up= sum(epa, na.rm = TRUE),
    
  EPA_passing_given = sum(
      epa[play_type == "pass"],
      na.rm = TRUE),
    
 
    EPA_rushing_given = sum(
      epa[play_type == "run"],
      na.rm = TRUE),
    
 
    QB_EPA_given = sum(qb_epa, na.rm = TRUE),
    
 
    defensive_plays = n(),
    
    pass_plays_faced = sum(play_type == "pass", na.rm = TRUE),
    
    rush_plays_faced = sum(play_type == "run", na.rm = TRUE),
    
 
    EPA_given_up_per_play =
      EPA_given_up / defensive_plays,
    
    EPA_passing_given_per_play =
      EPA_passing_given / pass_plays_faced,
    
    EPA_rushing_given_per_play =
      EPA_rushing_given / rush_plays_faced,
    
    QB_EPA_given_per_play =
      QB_EPA_given / pass_plays_faced,
    
    .groups = "drop")%>%
  rename(team = defteam)
```

Joined Data
```{r}
league_wk_2_summary = offense_wk_2%>%
  left_join(
    defense_wk_2,
    by = "team")%>%
  left_join(
    team_points_wk_2,
    by = "team")%>%
  mutate(
    Net_EPA =
      EPA_generated - EPA_given_up,
    
    Net_EPA_per_play =
      EPA_generated_per_play -
      EPA_given_up_per_play,
    
    turnover_differential =
      defensive_turnovers - offensive_turnovers)%>%
  mutate(
    points_rank = min_rank(desc(points)),
    
    offensive_plays_rank = min_rank(desc(offensive_plays)),
    
    defensive_plays_rank = min_rank((defensive_plays)),
    
    yards_rank = min_rank(desc(yards)),
    
    passing_yards_rank = min_rank(desc(passing_yards)),
    
    rushing_yards_rank = min_rank(desc(rushing_yards)),
    
    offensive_turnovers_rank = min_rank(offensive_turnovers),
    
    EPA_generated_rank = min_rank(desc(EPA_generated)),
    
    EPA_passing_rank = min_rank(desc(EPA_passing)),
    
    EPA_rushing_rank = min_rank(desc(EPA_rushing)),
    
    QB_EPA_rank = min_rank(desc(QB_EPA)),
    
    EPA_generated_per_play_rank = min_rank(desc(EPA_generated_per_play)),
    
    EPA_passing_per_play_rank = min_rank(desc(EPA_passing_per_play)),
    
    EPA_rushing_per_play_rank = min_rank(desc(EPA_rushing_per_play)),
    
    QB_EPA_per_play_rank = min_rank(desc(QB_EPA_per_play)),
    
    yards_allowed_rank = min_rank(yards_allowed),
    
    passing_yards_allowed_rank = min_rank((passing_yards_allowed)),
    
    rushing_yards_allowed_rank = min_rank((rushing_yards_allowed)),
    
    defensive_turnovers_rank = min_rank(desc(defensive_turnovers)),
    
    EPA_given_up_rank = min_rank(EPA_given_up),
    
    EPA_passing_given_rank = min_rank(EPA_passing_given),
    
    EPA_rushing_given_rank = min_rank(EPA_rushing_given),
    
    QB_EPA_given_rank = min_rank(QB_EPA_given),
    
    EPA_given_up_per_play_rank = min_rank(EPA_given_up_per_play),
    
    EPA_passing_given_per_play_rank = min_rank(EPA_passing_given_per_play),
     
    EPA_rushing_given_per_play_rank = min_rank(EPA_rushing_given_per_play),
    
    QB_EPA_given_per_play_rank = min_rank(QB_EPA_given_per_play),
    
    Net_EPA_rank = min_rank(desc(Net_EPA)),
    
    Net_EPA_per_play_rank = min_rank(desc(Net_EPA_per_play)),
    
    turnover_differential_rank = min_rank(desc(turnover_differential)))


```


Cleaning it up

```{r}
league_wk_2_summary = league_wk_2_summary %>%
  select(
    team, 
    points, 
    points_rank, 
    offensive_plays, 
    offensive_plays_rank, 
    yards, 
    yards_rank, 
    passing_yards, 
    passing_yards_rank, 
    rushing_yards, 
    rushing_yards_rank, 
    turnover_differential, 
    turnover_differential_rank, 
    offensive_turnovers, 
    offensive_turnovers_rank, 
    defensive_plays,
    yards_allowed,
    yards_allowed_rank,
    passing_yards_allowed,
    passing_yards_allowed_rank,
    rushing_yards_allowed,
    rushing_yards_allowed_rank,
    defensive_turnovers, 
    defensive_turnovers_rank, 
    Net_EPA, 
    Net_EPA_rank, 
    Net_EPA_per_play, 
    Net_EPA_per_play_rank, 
    EPA_generated, 
    EPA_generated_rank, 
    EPA_passing, 
    EPA_passing_rank, 
    EPA_rushing, 
    EPA_rushing_rank, 
    QB_EPA, 
    QB_EPA_rank, 
    EPA_passing_per_play, 
    EPA_passing_per_play_rank, 
    EPA_rushing_per_play, 
    EPA_rushing_per_play_rank, 
    QB_EPA_per_play, 
    QB_EPA_per_play_rank, 
    EPA_generated_per_play, 
    EPA_generated_per_play_rank,
    EPA_given_up, 
    EPA_given_up_rank,
    EPA_passing_given,
    EPA_passing_given_rank,
    EPA_rushing_given,
    EPA_rushing_given_rank,
    QB_EPA_given,
    QB_EPA_given_rank,
    EPA_given_up_per_play, 
    EPA_given_up_per_play_rank,
    EPA_passing_given_per_play,
    EPA_passing_given_per_play_rank,
    EPA_rushing_given_per_play,
    EPA_rushing_given_per_play_rank,
    QB_EPA_given_per_play,
    QB_EPA_given_per_play_rank)
```


Chargers Data Table

```{r}
Chargers_Week_2_Summary = league_wk_2_summary %>%
  filter(team == "LAC")

DT::datatable(Chargers_Week_2_Summary)
```


Comparing Chargers Week 2 Adv Statistics to the League

```{r}
stats_to_compare = c(
  "points",
  "offensive_plays",
  "yards",
  "passing_yards",
  "rushing_yards",
  "turnover_differential",
  "offensive_turnovers",
  "defensive_plays",
  "defensive_turnovers",
  "Net_EPA",
  "Net_EPA_per_play",
  "EPA_generated",
  "EPA_passing",
  "EPA_rushing",
  "QB_EPA",
  "EPA_generated_per_play",
  "EPA_passing_per_play",
  "EPA_rushing_per_play",
  "QB_EPA_per_play")

defensive_stats_to_compare = c(
  "EPA_given_up",
  "EPA_passing_given",
  "EPA_rushing_given",
  "QB_EPA_given",
  "EPA_given_up_per_play",
  "EPA_passing_given_per_play",
  "EPA_rushing_given_per_play",
  "QB_EPA_given_per_play")

Chargers_vs_League = league_wk_2_summary%>%
  filter(team == "LAC")%>%
  mutate(
    across(
      all_of(stats_to_compare),
      ~ .x - mean(league_wk_2_summary[[cur_column()]], na.rm = TRUE),
      .names = "{.col}_vs_league_avg"),
    
    across(
      all_of(defensive_stats_to_compare),
      ~ mean(league_wk_2_summary[[cur_column()]], na.rm = TRUE) - .x,
      .names = "{.col}_vs_league_avg"))
```

Data Table of Chargers week 1 stats
```{r}
DT::datatable(
  Chargers_vs_League,
  options = list(
    scrollX = TRUE,
    pageLength = 10))
```

Chargers vs League Stats Heatmap week 2:

Names
```{r}
heatmap_stats = c(
  "points",
  "yards",
  "passing_yards",
  "rushing_yards",
  "turnover_differential",
  "Net_EPA",
  "Net_EPA_per_play",
  "EPA_passing",
  "EPA_rushing",
  "QB_EPA",
  "EPA_passing_per_play",
  "EPA_rushing_per_play",
  "QB_EPA_per_play",
  "EPA_passing_given",
  "EPA_rushing_given",
  "QB_EPA_given",
  "EPA_passing_given_per_play",
  "EPA_rushing_given_per_play",
  "QB_EPA_given_per_play")
```

Offense
```{r}
heatmap_stats = c(
  "points",
  "yards",
  "passing_yards",
  "rushing_yards",
  "turnover_differential",
  "Net_EPA",
  "Net_EPA_per_play",
  "EPA_passing",
  "EPA_rushing",
  "QB_EPA",
  "EPA_passing_per_play",
  "EPA_rushing_per_play",
  "QB_EPA_per_play",
  "EPA_passing_given",
  "EPA_rushing_given",
  "QB_EPA_given",
  "EPA_passing_given_per_play",
  "EPA_rushing_given_per_play",
  "QB_EPA_given_per_play")

heatmap_data = league_wk_2_summary%>%
  select(team, all_of(heatmap_stats))%>%
  pivot_longer(
    cols = -team,
    names_to = "stat",
    values_to = "value")%>%
  group_by(stat)%>%
  mutate(
    league_avg = mean(value, na.rm = TRUE),
    vs_league_avg = value - league_avg)%>%
  ungroup()
```

Defense
```{r}
defensive_stats = c(
  "EPA_passing_given",
  "EPA_rushing_given",
  "QB_EPA_given",
  "EPA_passing_given_per_play",
  "EPA_rushing_given_per_play",
  "QB_EPA_given_per_play")

heatmap_data = heatmap_data%>%
  mutate(
    performance_vs_avg = case_when(
      stat %in% defensive_stats ~ -vs_league_avg,
      TRUE ~ vs_league_avg))

#Z score
heatmap_data = heatmap_data%>%
  group_by(stat)%>%
  mutate(
    z_score = (performance_vs_avg - mean(performance_vs_avg, na.rm = TRUE)) /
              sd(performance_vs_avg, na.rm = TRUE))%>%
  ungroup()
```

Heatmap
```{r}
ggplot(
  heatmap_data,
  aes(
    x = team,
    y = stat,
    fill = z_score))+
  geom_tile(color = "grey20")+
  
  scale_fill_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = 0)+
  
  labs(
    title = "NFL Week 2 Performance vs. League Average",
    subtitle = "Positive values indicate above-average performance",
    x = "Team",
    y = "Statistic",
    fill = "Vs. League Avg.")+
  
  theme_minimal()+
  theme(
    plot.background = element_rect(fill = "black", color = NA),
    panel.background = element_rect(fill = "black", color = NA),
    panel.grid = element_blank(),
    
    text = element_text(color = "white"),
    axis.text = element_text(color = "white"),
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      color = "white"),
    axis.title = element_text(color = "white"),
    
    plot.title = element_text(
      color = "white",
      face = "bold"),
    plot.subtitle = element_text(color = "grey80"),
    
    legend.background = element_rect(fill = "black", color = NA),
    legend.text = element_text(color = "white"),
    legend.title = element_text(color = "white"))
```

Comparing week 1 to week 2:

```{r}
league_wk_1 = league%>%
  filter(week == 1)

team_points_wk_1 = league_wk_1%>%
  group_by(game_id)%>%
  summarise(
    home_team = first(home_team),
    away_team = first(away_team),
    home_points = max(total_home_score, na.rm = TRUE),
    away_points = max(total_away_score, na.rm = TRUE),
    .groups = "drop")

team_points_wk_1 = bind_rows(
  team_points_wk_1 %>%
    transmute(
      team = home_team,
      points = home_points),
  
  team_points_wk_1 %>%
    transmute(
      team = away_team,
      points = away_points))
#Offense
offense_wk_1 = league_wk_1%>%
  filter(
    !is.na(posteam),
    !play_type %in% c(
      "kickoff",
      "extra_point",
      "field_goal"))%>%
  group_by(posteam)%>%
  summarise(
    yards = sum(yards_gained, na.rm = TRUE),
    
    passing_yards = sum(passing_yards, na.rm = TRUE),
    
    rushing_yards = sum(rushing_yards, na.rm = TRUE),
    
    offensive_turnovers = sum(
      interception + fumble_lost,
      na.rm = TRUE),
    

    EPA_generated = sum(epa, na.rm = TRUE),
    

    EPA_passing = sum(
      epa[play_type == "pass"],
      na.rm = TRUE),
    
 
    EPA_rushing = sum(
      epa[play_type == "run"],
      na.rm = TRUE),
    
 
    QB_EPA = sum(qb_epa, na.rm = TRUE),
    
 
    offensive_plays = n(),
    
    pass_plays = sum(play_type == "pass", na.rm = TRUE),
    
    rush_plays = sum(play_type == "run", na.rm = TRUE),
    
 
    EPA_generated_per_play =
      EPA_generated / offensive_plays,
    
    EPA_passing_per_play =
      EPA_passing / pass_plays,
    
    EPA_rushing_per_play =
      EPA_rushing / rush_plays,
    
    QB_EPA_per_play =
      QB_EPA / pass_plays,
    
    .groups = "drop")%>%
  rename(team = posteam)

#Defense
defense_wk_1 = league_wk_1%>%
  filter(
    !is.na(defteam),
    !play_type %in% c(
      "kickoff",
      "extra_point",
      "field_goal"))%>%
  group_by(defteam)%>%
  summarise(
    yards_allowed = sum(yards_gained, na.rm = TRUE),
    
    passing_yards_allowed = sum(passing_yards, na.rm = TRUE),
    
    rushing_yards_allowed = sum(rushing_yards, na.rm = TRUE),
    
    defensive_turnovers = sum(
  interception + fumble_forced,
  na.rm = TRUE),
  
  EPA_given_up= sum(epa, na.rm = TRUE),
    
  EPA_passing_given = sum(
      epa[play_type == "pass"],
      na.rm = TRUE),
    
 
    EPA_rushing_given = sum(
      epa[play_type == "run"],
      na.rm = TRUE),
    
 
    QB_EPA_given = sum(qb_epa, na.rm = TRUE),
    
 
    defensive_plays = n(),
    
    pass_plays_faced = sum(play_type == "pass", na.rm = TRUE),
    
    rush_plays_faced = sum(play_type == "run", na.rm = TRUE),
    
 
    EPA_given_up_per_play =
      EPA_given_up / defensive_plays,
    
    EPA_passing_given_per_play =
      EPA_passing_given / pass_plays_faced,
    
    EPA_rushing_given_per_play =
      EPA_rushing_given / rush_plays_faced,
    
    QB_EPA_given_per_play =
      QB_EPA_given / pass_plays_faced,
    
    .groups = "drop")%>%
  rename(team = defteam)


```

Joined Data
```{r}
league_wk_1_summary = offense_wk_1%>%
  left_join(
    defense_wk_1,
    by = "team")%>%
  left_join(
    team_points_wk_1,
    by = "team")%>%
  mutate(
    Net_EPA =
      EPA_generated - EPA_given_up,
    
    Net_EPA_per_play =
      EPA_generated_per_play -
      EPA_given_up_per_play,
    
    turnover_differential =
      defensive_turnovers - offensive_turnovers)%>%
  mutate(
    points_rank = min_rank(desc(points)),
    
    offensive_plays_rank = min_rank(desc(offensive_plays)),
    
    defensive_plays_rank = min_rank((defensive_plays)),
    
    yards_rank = min_rank(desc(yards)),
    
    passing_yards_rank = min_rank(desc(passing_yards)),
    
    rushing_yards_rank = min_rank(desc(rushing_yards)),
    
    offensive_turnovers_rank = min_rank(offensive_turnovers),
    
    EPA_generated_rank = min_rank(desc(EPA_generated)),
    
    EPA_passing_rank = min_rank(desc(EPA_passing)),
    
    EPA_rushing_rank = min_rank(desc(EPA_rushing)),
    
    QB_EPA_rank = min_rank(desc(QB_EPA)),
    
    EPA_generated_per_play_rank = min_rank(desc(EPA_generated_per_play)),
    
    EPA_passing_per_play_rank = min_rank(desc(EPA_passing_per_play)),
    
    EPA_rushing_per_play_rank = min_rank(desc(EPA_rushing_per_play)),
    
    QB_EPA_per_play_rank = min_rank(desc(QB_EPA_per_play)),
    
    yards_allowed_rank = min_rank(yards_allowed),
    
    passing_yards_allowed_rank = min_rank((passing_yards_allowed)),
    
    rushing_yards_allowed_rank = min_rank((rushing_yards_allowed)),
    
    defensive_turnovers_rank = min_rank(desc(defensive_turnovers)),
    
    EPA_given_up_rank = min_rank(EPA_given_up),
    
    EPA_passing_given_rank = min_rank(EPA_passing_given),
    
    EPA_rushing_given_rank = min_rank(EPA_rushing_given),
    
    QB_EPA_given_rank = min_rank(QB_EPA_given),
    
    EPA_given_up_per_play_rank = min_rank(EPA_given_up_per_play),
    
    EPA_passing_given_per_play_rank = min_rank(EPA_passing_given_per_play),
     
    EPA_rushing_given_per_play_rank = min_rank(EPA_rushing_given_per_play),
    
    QB_EPA_given_per_play_rank = min_rank(QB_EPA_given_per_play),
    
    Net_EPA_rank = min_rank(desc(Net_EPA)),
    
    Net_EPA_per_play_rank = min_rank(desc(Net_EPA_per_play)),
    
    turnover_differential_rank = min_rank(desc(turnover_differential)))


```


Cleaning it up

```{r}
league_wk_1_summary = league_wk_1_summary%>%
  select(
    team, 
    points, 
    points_rank, 
    offensive_plays, 
    offensive_plays_rank, 
    yards, yards_rank, 
    passing_yards, 
    passing_yards_rank, 
    rushing_yards, 
    rushing_yards_rank, 
    turnover_differential, 
    turnover_differential_rank, 
    offensive_turnovers, 
    offensive_turnovers_rank, 
    defensive_plays,
    yards_allowed,
    passing_yards_allowed,
    rushing_yards_allowed,
    defensive_turnovers, 
    defensive_turnovers_rank, 
    Net_EPA, 
    Net_EPA_rank, 
    Net_EPA_per_play, 
    Net_EPA_per_play_rank, 
    EPA_generated, 
    EPA_generated_rank, 
    EPA_passing, 
    EPA_passing_rank, 
    EPA_rushing, 
    EPA_rushing_rank, 
    QB_EPA, 
    QB_EPA_rank, 
    EPA_passing_per_play, 
    EPA_passing_per_play_rank, 
    EPA_rushing_per_play, 
    EPA_rushing_per_play_rank, 
    QB_EPA_per_play, 
    QB_EPA_per_play_rank, 
    EPA_generated_per_play, 
    EPA_generated_per_play_rank,
    EPA_given_up, 
    EPA_given_up_rank,
    EPA_passing_given,
    EPA_passing_given_rank,
    EPA_rushing_given,
    EPA_rushing_given_rank,
    QB_EPA_given,
    QB_EPA_given_rank,
    EPA_given_up_per_play, 
    EPA_given_up_per_play_rank,
    EPA_passing_given_per_play,
    EPA_passing_given_per_play_rank,
    EPA_rushing_given_per_play,
    EPA_rushing_given_per_play_rank,
    QB_EPA_given_per_play,
    QB_EPA_given_per_play_rank
  )
```

Joining with Week 2 Summary

```{r}
offensive_stats = c(
  "points",
  "offensive_plays",
  "yards",
  "passing_yards",
  "rushing_yards",
  "turnover_differential",
  "defensive_turnovers",
  "Net_EPA",
  "Net_EPA_per_play",
  "EPA_generated",
  "EPA_passing",
  "EPA_rushing",
  "QB_EPA",
  "EPA_generated_per_play",
  "EPA_passing_per_play",
  "EPA_rushing_per_play",
  "QB_EPA_per_play")

defensive_stats = c(
  "defensive_plays",
  "yards_allowed",
  "passing_yards_allowed",
  "rushing_yards_allowed",
  "offensive_turnovers",
  "EPA_given_up",
  "EPA_passing_given",
  "EPA_rushing_given",
  "QB_EPA_given",
  "EPA_given_up_per_play",
  "EPA_passing_given_per_play",
  "EPA_rushing_given_per_play",
  "QB_EPA_given_per_play")


Chargers_Wk1_vs_Wk2 = league_wk_1_summary%>%
  filter(team == "LAC")%>%
  select(
    team,
    all_of(c(offensive_stats, defensive_stats)))%>%
  rename_with(
    ~ paste0(.x, "_wk1"),
    -team)%>%
  left_join(
    league_wk_2_summary%>%
      filter(team == "LAC")%>%
      select(
        team,
        all_of(c(offensive_stats, defensive_stats)))%>%
      rename_with(
        ~ paste0(.x, "_wk2"),
        -team),
    by = "team")%>%
mutate(
  # Higher is better
  points_change_from_week_1 =
    points_wk2 - points_wk1,

  offensive_plays_change_from_week_1 =
    offensive_plays_wk2 - offensive_plays_wk1,

  yards_change_from_week_1 =
    yards_wk2 - yards_wk1,

  passing_yards_change_from_week_1 =
    passing_yards_wk2 - passing_yards_wk1,

  rushing_yards_change_from_week_1 =
    rushing_yards_wk2 - rushing_yards_wk1,

  turnover_differential_change_from_week_1 =
    turnover_differential_wk2 - turnover_differential_wk1,

  defensive_turnovers_change_from_week_1 =
    defensive_turnovers_wk2 - defensive_turnovers_wk1,

  Net_EPA_change_from_week_1 =
    Net_EPA_wk2 - Net_EPA_wk1,

  Net_EPA_per_play_change_from_week_1 =
    Net_EPA_per_play_wk2 - Net_EPA_per_play_wk1,

  EPA_generated_change_from_week_1 =
    EPA_generated_wk2 - EPA_generated_wk1,

  EPA_passing_change_from_week_1 =
    EPA_passing_wk2 - EPA_passing_wk1,

  EPA_rushing_change_from_week_1 =
    EPA_rushing_wk2 - EPA_rushing_wk1,

  QB_EPA_change_from_week_1 =
    QB_EPA_wk2 - QB_EPA_wk1,

  EPA_generated_per_play_change_from_week_1 =
    EPA_generated_per_play_wk2 - EPA_generated_per_play_wk1,

  EPA_passing_per_play_change_from_week_1 =
    EPA_passing_per_play_wk2 - EPA_passing_per_play_wk1,

  EPA_rushing_per_play_change_from_week_1 =
    EPA_rushing_per_play_wk2 - EPA_rushing_per_play_wk1,

  QB_EPA_per_play_change_from_week_1 =
    QB_EPA_per_play_wk2 - QB_EPA_per_play_wk1,


  # Volume — simply Week 2 minus Week 1
  defensive_plays_change_from_week_1 =
    defensive_plays_wk2 - defensive_plays_wk1,


  # Lower is better — reverse the subtraction
  yards_allowed_change_from_week_1 =
    yards_allowed_wk1 - yards_allowed_wk2,

  passing_yards_allowed_change_from_week_1 =
    passing_yards_allowed_wk1 - passing_yards_allowed_wk2,

  rushing_yards_allowed_change_from_week_1 =
    rushing_yards_allowed_wk1 - rushing_yards_allowed_wk2,

  offensive_turnovers_change_from_week_1 =
    offensive_turnovers_wk1 - offensive_turnovers_wk2,

  EPA_given_up_change_from_week_1 =
    EPA_given_up_wk1 - EPA_given_up_wk2,

  EPA_passing_given_change_from_week_1 =
    EPA_passing_given_wk1 - EPA_passing_given_wk2,

  EPA_rushing_given_change_from_week_1 =
    EPA_rushing_given_wk1 - EPA_rushing_given_wk2,

  QB_EPA_given_change_from_week_1 =
    QB_EPA_given_wk1 - QB_EPA_given_wk2,

  EPA_given_up_per_play_change_from_week_1 =
    EPA_given_up_per_play_wk1 - EPA_given_up_per_play_wk2,

  EPA_passing_given_per_play_change_from_week_1 =
    EPA_passing_given_per_play_wk1 - EPA_passing_given_per_play_wk2,

  EPA_rushing_given_per_play_change_from_week_1 =
    EPA_rushing_given_per_play_wk1 - EPA_rushing_given_per_play_wk2,

  QB_EPA_given_per_play_change_from_week_1 =
    QB_EPA_given_per_play_wk1 - QB_EPA_given_per_play_wk2)
```


#Week 1 vs Week 2 Data Table:
```{r}
DT::datatable(
  Chargers_Wk1_vs_Wk2,
  options = list(
    scrollX = TRUE,
    pageLength = 10))
```

Week 2/Week 1 Difference HeatMap Data

```{r}
Chargers_Wk1_Wk2_table = tibble(
  
  Stat = c(
    # Offense
    "Points",
    "Yards",
    "Passing Yards",
    "Rushing Yards",
    "Turnover Differential",
    "Defensive Turnovers",
    "Net EPA",
    "Net EPA / Play",
    "EPA Generated",
    "EPA Passing",
    "EPA Rushing",
    "QB EPA",
    "EPA Generated / Play",
    "EPA Passing / Play",
    "EPA Rushing / Play",
    "QB EPA / Play",
    
    # Defense
    "Yards Allowed",
    "Passing Yards Allowed",
    "Rushing Yards Allowed",
    "Offensive Turnovers",
    "EPA Given Up",
    "EPA Passing Given",
    "EPA Rushing Given",
    "QB EPA Given",
    "EPA Given Up / Play",
    "EPA Passing Given / Play",
    "EPA Rushing Given / Play",
    "QB EPA Given / Play"),
  
  Week_1 = c(
    Chargers_Wk1_vs_Wk2$points_wk1,
    Chargers_Wk1_vs_Wk2$yards_wk1,
    Chargers_Wk1_vs_Wk2$passing_yards_wk1,
    Chargers_Wk1_vs_Wk2$rushing_yards_wk1,
    Chargers_Wk1_vs_Wk2$turnover_differential_wk1,
    Chargers_Wk1_vs_Wk2$defensive_turnovers_wk1,
    Chargers_Wk1_vs_Wk2$Net_EPA_wk1,
    Chargers_Wk1_vs_Wk2$Net_EPA_per_play_wk1,
    Chargers_Wk1_vs_Wk2$EPA_generated_wk1,
    Chargers_Wk1_vs_Wk2$EPA_passing_wk1,
    Chargers_Wk1_vs_Wk2$EPA_rushing_wk1,
    Chargers_Wk1_vs_Wk2$QB_EPA_wk1,
    Chargers_Wk1_vs_Wk2$EPA_generated_per_play_wk1,
    Chargers_Wk1_vs_Wk2$EPA_passing_per_play_wk1,
    Chargers_Wk1_vs_Wk2$EPA_rushing_per_play_wk1,
    Chargers_Wk1_vs_Wk2$QB_EPA_per_play_wk1,
    
    Chargers_Wk1_vs_Wk2$yards_allowed_wk1,
    Chargers_Wk1_vs_Wk2$passing_yards_allowed_wk1,
    Chargers_Wk1_vs_Wk2$rushing_yards_allowed_wk1,
    Chargers_Wk1_vs_Wk2$offensive_turnovers_wk1,
    Chargers_Wk1_vs_Wk2$EPA_given_up_wk1,
    Chargers_Wk1_vs_Wk2$EPA_passing_given_wk1,
    Chargers_Wk1_vs_Wk2$EPA_rushing_given_wk1,
    Chargers_Wk1_vs_Wk2$QB_EPA_given_wk1,
    Chargers_Wk1_vs_Wk2$EPA_given_up_per_play_wk1,
    Chargers_Wk1_vs_Wk2$EPA_passing_given_per_play_wk1,
    Chargers_Wk1_vs_Wk2$EPA_rushing_given_per_play_wk1,
    Chargers_Wk1_vs_Wk2$QB_EPA_given_per_play_wk1),
  
  Week_2 = c(
    Chargers_Wk1_vs_Wk2$points_wk2,
    Chargers_Wk1_vs_Wk2$yards_wk2,
    Chargers_Wk1_vs_Wk2$passing_yards_wk2,
    Chargers_Wk1_vs_Wk2$rushing_yards_wk2,
    Chargers_Wk1_vs_Wk2$turnover_differential_wk2,
    Chargers_Wk1_vs_Wk2$defensive_turnovers_wk2,
    Chargers_Wk1_vs_Wk2$Net_EPA_wk2,
    Chargers_Wk1_vs_Wk2$Net_EPA_per_play_wk2,
    Chargers_Wk1_vs_Wk2$EPA_generated_wk2,
    Chargers_Wk1_vs_Wk2$EPA_passing_wk2,
    Chargers_Wk1_vs_Wk2$EPA_rushing_wk2,
    Chargers_Wk1_vs_Wk2$QB_EPA_wk2,
    Chargers_Wk1_vs_Wk2$EPA_generated_per_play_wk2,
    Chargers_Wk1_vs_Wk2$EPA_passing_per_play_wk2,
    Chargers_Wk1_vs_Wk2$EPA_rushing_per_play_wk2,
    Chargers_Wk1_vs_Wk2$QB_EPA_per_play_wk2,
    
    Chargers_Wk1_vs_Wk2$yards_allowed_wk2,
    Chargers_Wk1_vs_Wk2$passing_yards_allowed_wk2,
    Chargers_Wk1_vs_Wk2$rushing_yards_allowed_wk2,
    Chargers_Wk1_vs_Wk2$offensive_turnovers_wk2,
    Chargers_Wk1_vs_Wk2$EPA_given_up_wk2,
    Chargers_Wk1_vs_Wk2$EPA_passing_given_wk2,
    Chargers_Wk1_vs_Wk2$EPA_rushing_given_wk2,
    Chargers_Wk1_vs_Wk2$QB_EPA_given_wk2,
    Chargers_Wk1_vs_Wk2$EPA_given_up_per_play_wk2,
    Chargers_Wk1_vs_Wk2$EPA_passing_given_per_play_wk2,
    Chargers_Wk1_vs_Wk2$EPA_rushing_given_per_play_wk2,
    Chargers_Wk1_vs_Wk2$QB_EPA_given_per_play_wk2),
  
  Difference = c(
    Chargers_Wk1_vs_Wk2$points_change_from_week_1,
    Chargers_Wk1_vs_Wk2$yards_change_from_week_1,
    Chargers_Wk1_vs_Wk2$passing_yards_change_from_week_1,
    Chargers_Wk1_vs_Wk2$rushing_yards_change_from_week_1,
    Chargers_Wk1_vs_Wk2$turnover_differential_change_from_week_1,
    Chargers_Wk1_vs_Wk2$defensive_turnovers_change_from_week_1,
    Chargers_Wk1_vs_Wk2$Net_EPA_change_from_week_1,
    Chargers_Wk1_vs_Wk2$Net_EPA_per_play_change_from_week_1,
    Chargers_Wk1_vs_Wk2$EPA_generated_change_from_week_1,
    Chargers_Wk1_vs_Wk2$EPA_passing_change_from_week_1,
    Chargers_Wk1_vs_Wk2$EPA_rushing_change_from_week_1,
    Chargers_Wk1_vs_Wk2$QB_EPA_change_from_week_1,
    Chargers_Wk1_vs_Wk2$EPA_generated_per_play_change_from_week_1,
    Chargers_Wk1_vs_Wk2$EPA_passing_per_play_change_from_week_1,
    Chargers_Wk1_vs_Wk2$EPA_rushing_per_play_change_from_week_1,
    Chargers_Wk1_vs_Wk2$QB_EPA_per_play_change_from_week_1,
    
    Chargers_Wk1_vs_Wk2$yards_allowed_change_from_week_1,
    Chargers_Wk1_vs_Wk2$passing_yards_allowed_change_from_week_1,
    Chargers_Wk1_vs_Wk2$rushing_yards_allowed_change_from_week_1,
    Chargers_Wk1_vs_Wk2$offensive_turnovers_change_from_week_1,
    Chargers_Wk1_vs_Wk2$EPA_given_up_change_from_week_1,
    Chargers_Wk1_vs_Wk2$EPA_passing_given_change_from_week_1,
    Chargers_Wk1_vs_Wk2$EPA_rushing_given_change_from_week_1,
    Chargers_Wk1_vs_Wk2$QB_EPA_given_change_from_week_1,
    Chargers_Wk1_vs_Wk2$EPA_given_up_per_play_change_from_week_1,
    Chargers_Wk1_vs_Wk2$EPA_passing_given_per_play_change_from_week_1,
    Chargers_Wk1_vs_Wk2$EPA_rushing_given_per_play_change_from_week_1,
    Chargers_Wk1_vs_Wk2$QB_EPA_given_per_play_change_from_week_1)) %>%
  mutate(
    Direction = case_when(
      Difference > 0 ~ "↑",
      Difference < 0 ~ "↓",
      TRUE ~ "→"))
```

Heatmap
```{r}
Chargers_Wk1_Wk2_table%>%
  gt()%>%
  tab_header(
    title = "Chargers: Week 1 vs. Week 2",
    subtitle = "Positive differences indicate improved performance")%>%
  cols_label(
    Stat = "Statistic",
    Week_1 = "Week 1",
    Week_2 = "Week 2",
    Difference = "Difference",
    Direction = "Direction")%>%
  fmt_number(
    columns = c(Week_1, Week_2, Difference),
    decimals = 2)%>%
  cols_align(
    align = "center",
    columns = c(Week_1, Week_2, Difference, Direction))%>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(everything()))%>%
  tab_style(
    style = cell_text(size = px(20), weight = "bold"),
    locations = cells_body(columns = Direction))%>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = Stat))%>%
  opt_row_striping()%>%
  opt_table_font(
    font = list(
      google_font("Roboto"),
      default_fonts()))%>%
  tab_options(
    table.font.size = px(14),
    heading.title.font.size = px(22),
    heading.subtitle.font.size = px(14))
```




