---
title: "Statistically Speaking 2026 Week 2"
author: "Josh Garzaniti"
date: "2026-09-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install and load in needed libraries
```{r}
library(nflfastR)
library(nflreadr)
library(tidyverse)
library(stringr)
library(ggplot2)
library(ggdark)
library(caret)
library(xgboost)
library(randomForest)
library(ggrepel)
library(nflplotR)
library(mice)
library(lme4)
library(here)
library(DT)
library(htmlwidgets)
library(gt)
```
Load in our play by play data
```{r}
pbp = load_pbp(2026)

#Create redzone variable:
pbp = pbp%>%
  mutate(redzone = ifelse(yardline_100 <= 20, 1, 0))
```

First I want to separate into 2 dataframes: Chargers and the rest of the league

```{r}
chargers = pbp%>%
  filter(home_team == "LAC" | away_team == "LAC")

league = pbp

chargers_offense = chargers%>%
  filter(posteam == "LAC" & !play_type %in% c("kickoff", "extra_point", "field_goal"))

chargers_defense = chargers%>%
  filter(defteam == "LAC" & !play_type %in% c("kickoff", "extra_point", "field_goal"))

chargers_st = chargers%>%
  filter(play_type %in% c("kickoff", "extra_point", "field_goal"))
```


Week 3 vs Buffalo Bills

```{r}
chargers_wk_3 = chargers%>%
  filter(posteam == "LAC" 
         & week == 3)

chargers_offense_wk_3 = chargers%>%
  filter(posteam == "LAC" 
         & week == 3
         & !play_type %in% c("kickoff", "extra_point", "field_goal"))

chargers_defense_wk_3 = chargers%>%
  filter(defteam == "LAC"
         & week == 3
         & !play_type %in% c("kickoff", "extra_point", "field_goal"))

chargers_st_wk_3 = chargers%>%
  filter(play_type %in% c("kickoff", "extra_point", "field_goal")
         & week == 3)
```


Summary:

```{r}
total_wk3_points = max(chargers$total_home_score, na.rm = TRUE)

total_wk3_points
  
total_wk3_points_allowed = max(chargers$total_away_score, na.rm = TRUE)

total_wk3_points_allowed

total_wk3_yards = sum(chargers_offense_wk_3$yards_gained, na.rm = TRUE)

total_wk3_yards

total_wk3_yards_allowed = sum(chargers_defense_wk_3$yards_gained, na.rm = TRUE)

total_wk3_yards_allowed

total_wk3_turnover_dif = (sum(chargers_defense_wk_3$interception, chargers_defense_wk_3$fumble_forced, na.rm = TRUE) - sum(chargers_offense_wk_3$interception, chargers_offense_wk_3$fumble_lost, na.rm = TRUE))

total_wk3_turnover_dif

total_wk3_EPA_generated = sum(chargers_offense_wk_3$epa, na.rm = TRUE)

total_wk3_EPA_generated

total_wk3_EPA_given_up = sum(chargers_defense$epa, na.rm = TRUE)

total_wk3_EPA_given_up

total_wk3_Net_EPA = total_wk3_EPA_generated - total_wk3_EPA_given_up

total_wk3_Net_EPA

total_wk3_EPA_generated_per_play = total_wk3_EPA_generated/nrow(chargers_offense_wk_3)

total_wk3_EPA_generated_per_play

total_wk3_EPA_given_up_per_play = total_wk3_EPA_given_up/nrow(chargers_defense_wk_3)

total_wk3_EPA_given_up_per_play

total_wk3_Net_EPA_per_play = total_wk3_EPA_generated_per_play - total_wk3_EPA_given_up_per_play

total_wk3_Net_EPA_per_play

```

Chargers 14
Raiders 26
Chargers yards: 310
Raiders yards: 305
Turnover differential: -1
EPA generated: -19.12942
EPA allowed: -5.740576
Net EPA: -13.38885
EPA generated per play: -0.2484341
EPA allowed per play:-0.07863802
Net EPA per play: -0.1697961


## How did the league do?

```{r}
league_wk_3 = league%>%
  filter(week == 3)

```

Generate Stats for each team:

Scoring Week 3
```{r}
team_points_wk_3 = league_wk_3%>%
  group_by(game_id)%>%
  summarise(
    home_team = first(home_team),
    away_team = first(away_team),
    home_points = max(total_home_score, na.rm = TRUE),
    away_points = max(total_away_score, na.rm = TRUE),
    .groups = "drop")

team_points_wk_3 = bind_rows(
  team_points_wk_3%>%
    transmute(
      team = home_team,
      points = home_points),
  
  team_points_wk_3%>%
    transmute(
      team = away_team,
      points = away_points))
```

Offensive Stats
```{r}
offense_wk_3 = league_wk_3%>%
  filter(
    !is.na(posteam),
    !play_type %in% c(
      "kickoff",
      "extra_point",
      "field_goal"))%>%
  group_by(posteam)%>%
  summarise(
    
    yards = sum(yards_gained, na.rm = TRUE),
    
    passing_yards = sum(passing_yards, na.rm = TRUE),
    
    rushing_yards = sum(rushing_yards, na.rm = TRUE),
    
    offensive_turnovers = sum(
      interception + fumble_lost,
      na.rm = TRUE),
    
    EPA_generated = sum(epa, na.rm = TRUE),
    
    EPA_passing = sum(
      epa[play_type == "pass"],
      na.rm = TRUE),
    
    EPA_rushing = sum(
      epa[play_type == "run"],
      na.rm = TRUE),
    
    QB_EPA = sum(qb_epa, na.rm = TRUE),
    
    redzone_EPA = sum(
      epa[redzone == 1],
      na.rm = TRUE),
    
    redzone_drives = n_distinct(
      drive[redzone == 1]),
    
    redzone_touchdowns = n_distinct(
      drive[redzone == 1 & touchdown == 1]),
    
    redzone_conversion_rate =
      redzone_touchdowns / redzone_drives,
    
    sacks_allowed = sum(sack, na.rm = TRUE),
    
    pressures_allowed = sum(qb_hit, na.rm = TRUE),
    
    incomplete_passes = sum(incomplete_pass, na.rm = TRUE),
    
    offensive_plays = n(),
    
    pass_plays = sum(
      play_type == "pass",
      na.rm = TRUE),
    
    rush_plays = sum(
      play_type == "run",
      na.rm = TRUE),
    
    sack_rate =
      sacks_allowed / pass_plays,
    
    pressure_allowed_rate = pressures_allowed / pass_plays,
    
    incomplete_pass_rate =
      incomplete_passes / pass_plays,
    
    third_downs_faced =
      sum(down == 3, na.rm = TRUE),
    
    third_down_conversion_rate =
      sum(third_down_converted, na.rm = TRUE) /
      third_downs_faced,
    
    EPA_generated_per_play =
      EPA_generated / offensive_plays,
    
    EPA_passing_per_play =
      EPA_passing / pass_plays,
    
    EPA_rushing_per_play =
      EPA_rushing / rush_plays,
    
    QB_EPA_per_play =
      QB_EPA / pass_plays,
    
    off_success_rate =
      sum(success, na.rm = TRUE) /
      offensive_plays,
    
    off_rush_success_rate =
      sum(success[play_type == "run"], na.rm = TRUE) /
      rush_plays,
    
    off_pass_success_rate =
      sum(success[play_type == "pass"], na.rm = TRUE) /
      pass_plays,
    
    off_average_air_yards_depth =
      mean(air_yards, na.rm = TRUE),
    
    .groups = "drop")%>%
  rename(team = posteam)
```

Defensive Stats
```{r}
defense_wk_3 = league_wk_3%>%
  filter(
    !is.na(defteam),
    !play_type %in% c(
      "kickoff",
      "extra_point",
      "field_goal"))%>%
  group_by(defteam)%>%
  summarise(
    
    yards_allowed = sum(
      yards_gained,
      na.rm = TRUE),
    
    passing_yards_allowed = sum(
      passing_yards,
      na.rm = TRUE),
    
    rushing_yards_allowed = sum(
      rushing_yards,
      na.rm = TRUE),
    
    defensive_takeaways = sum(
      interception + fumble_forced,
      na.rm = TRUE),
    
    EPA_given_up = sum(
      epa,
      na.rm = TRUE),
    
    EPA_passing_given = sum(
      epa[play_type == "pass"],
      na.rm = TRUE),
    
    EPA_rushing_given = sum(
      epa[play_type == "run"],
      na.rm = TRUE),
    
    QB_EPA_given = sum(
      qb_epa,
      na.rm = TRUE),
    
    redzone_EPA_given = sum(
      epa[redzone == 1],
      na.rm = TRUE),
    
    redzone_drives_allowed = n_distinct(
      drive[redzone == 1]),
    
    redzone_touchdowns_allowed = n_distinct(
      drive[redzone == 1 & touchdown == 1]),
    
    redzone_conversion_rate_allowed =
      redzone_touchdowns_allowed /
      redzone_drives_allowed,
    
    sacks_forced = sum(
      sack,
      na.rm = TRUE),
    
    pressures = sum(qb_hit, na.rm = TRUE),
    
    incomplete_passes_forced = sum(
      incomplete_pass,
      na.rm = TRUE),
    
    defensive_plays = n(),
    
    pass_plays_faced = sum(
      play_type == "pass",
      na.rm = TRUE),
    
    rush_plays_faced = sum(
      play_type == "run",
      na.rm = TRUE),
    
    sack_forced_rate =
      sacks_forced / pass_plays_faced,
    
    pressure_rate = pressures / pass_plays_faced, 
    
    incomplete_pass_forced_rate =
      incomplete_passes_forced /
      pass_plays_faced,
    
    third_downs_forced =
      sum(down == 3, na.rm = TRUE),
    
    third_down_conversion_rate_allowed =
      sum(third_down_converted, na.rm = TRUE) /
      third_downs_forced,
    
    EPA_given_up_per_play =
      EPA_given_up / defensive_plays,
    
    EPA_passing_given_per_play =
      EPA_passing_given / pass_plays_faced,
    
    EPA_rushing_given_per_play =
      EPA_rushing_given / rush_plays_faced,
    
    QB_EPA_given_per_play =
      QB_EPA_given / pass_plays_faced,
    
    def_success_rate =
      1 - (
        sum(success, na.rm = TRUE) /
        defensive_plays),
    
    def_rush_success_rate =
      1 - (
        sum(success[play_type == "run"], na.rm = TRUE) /
        rush_plays_faced),
    
    def_pass_success_rate =
      1 - (
        sum(success[play_type == "pass"], na.rm = TRUE) /
        pass_plays_faced),
    
    def_average_air_yards_depth =
      mean(air_yards, na.rm = TRUE),
    
    .groups = "drop")%>%
  rename(team = defteam)
```

Joined Data
```{r}
league_wk_3_summary = offense_wk_3%>%
  left_join(
    defense_wk_3,
    by = "team")%>%
  left_join(
    team_points_wk_3,
    by = "team")%>%
  mutate(
    Net_EPA =
      EPA_generated - EPA_given_up,
    
    Net_EPA_per_play =
      EPA_generated_per_play -
      EPA_given_up_per_play,
    
    turnover_differential =
      defensive_takeaways - offensive_turnovers)%>%
  mutate(
    points_rank = min_rank(desc(points)),
    
    offensive_plays_rank = min_rank(desc(offensive_plays)),
    
    defensive_plays_rank = min_rank((defensive_plays)),
    
    yards_rank = min_rank(desc(yards)),
    
    passing_yards_rank = min_rank(desc(passing_yards)),
    
    rushing_yards_rank = min_rank(desc(rushing_yards)),
    
    offensive_turnovers_rank = min_rank(offensive_turnovers),
    
    EPA_generated_rank = min_rank(desc(EPA_generated)),
    
    EPA_passing_rank = min_rank(desc(EPA_passing)),
    
    EPA_rushing_rank = min_rank(desc(EPA_rushing)),
    
    QB_EPA_rank = min_rank(desc(QB_EPA)),
    
    redzone_EPA_rank = min_rank(desc(redzone_EPA)),
    
    redzone_conversion_rate_rank = min_rank(desc(redzone_conversion_rate)),
    
    sacks_allowed_rank = min_rank(sacks_allowed),
    
    sack_rate_rank = min_rank(sack_rate),
    
    pressures_allowed_rank = min_rank(pressures_allowed),
    
    pressure_allowed_rate_rank = min_rank(pressure_allowed_rate),
    
    incomplete_pass_rate_rank = min_rank(incomplete_pass_rate),
    
    third_down_conversion_rate_rank = min_rank(desc(third_down_conversion_rate)),
    
    EPA_generated_per_play_rank = min_rank(desc(EPA_generated_per_play)),
    
    EPA_passing_per_play_rank = min_rank(desc(EPA_passing_per_play)),
    
    EPA_rushing_per_play_rank = min_rank(desc(EPA_rushing_per_play)),
    
    QB_EPA_per_play_rank = min_rank(desc(QB_EPA_per_play)),
    
    off_success_rate_rank = min_rank(desc(off_success_rate)),
    
    off_rush_success_rate_rank = min_rank(desc(off_rush_success_rate)),
    
    off_pass_success_rate_rank = min_rank(desc(off_pass_success_rate)),
    
    off_average_air_yards_depth_rank = min_rank(desc(off_average_air_yards_depth)),
    
    yards_allowed_rank = min_rank(yards_allowed),
    
    passing_yards_allowed_rank = min_rank((passing_yards_allowed)),
    
    rushing_yards_allowed_rank = min_rank((rushing_yards_allowed)),
    
    defensive_takeaways_rank = min_rank(desc(defensive_takeaways)),
    
    EPA_given_up_rank = min_rank(EPA_given_up),
    
    EPA_passing_given_rank = min_rank(EPA_passing_given),
    
    EPA_rushing_given_rank = min_rank(EPA_rushing_given),
    
    QB_EPA_given_rank = min_rank(QB_EPA_given),
    
    redzone_EPA_given_rank = min_rank(redzone_EPA_given),
    
    redzone_conversion_rate_allowed_rank = min_rank(redzone_conversion_rate_allowed),
    
    sacks_forced_rank = min_rank(desc(sacks_forced)),
    
    sack_forced_rate_rank = min_rank(desc(sack_forced_rate)),
    
    pressures_rank = min_rank(desc(pressures)),
    
    pressure_rate_rank = min_rank(desc(pressure_rate)),
    
    incomplete_pass_forced_rate_rank = min_rank(desc(incomplete_pass_forced_rate)),
    
    third_down_conversion_rate_allowed_rank = min_rank(desc(third_down_conversion_rate_allowed)),
    
    EPA_given_up_per_play_rank = min_rank(EPA_given_up_per_play),
    
    EPA_passing_given_per_play_rank = min_rank(EPA_passing_given_per_play),
     
    EPA_rushing_given_per_play_rank = min_rank(EPA_rushing_given_per_play),
    
    QB_EPA_given_per_play_rank = min_rank(QB_EPA_given_per_play),
    
    def_success_rate_rank = min_rank(desc(def_success_rate)),
    
    def_rush_success_rate_rank = min_rank(desc(def_rush_success_rate)),
    
    def_pass_success_rate_rank = min_rank(desc(def_pass_success_rate)),
    
    def_average_air_yards_depth_rank = min_rank((def_average_air_yards_depth)),
    
    Net_EPA_rank = min_rank(desc(Net_EPA)),
    
    Net_EPA_per_play_rank = min_rank(desc(Net_EPA_per_play)),
    
    turnover_differential_rank = min_rank(desc(turnover_differential)))


```


Cleaning it up

```{r}
league_wk_3_summary = league_wk_3_summary%>%
  select(team, 
         points, 
         points_rank, 
         offensive_plays, 
         offensive_plays_rank, 
         yards, yards_rank, 
         passing_yards, 
         passing_yards_rank, 
         rushing_yards, 
         rushing_yards_rank, 
         turnover_differential, 
         turnover_differential_rank, 
         offensive_turnovers, 
         offensive_turnovers_rank, 
         defensive_plays, 
         defensive_takeaways, 
         defensive_takeaways_rank, 
         Net_EPA, 
         Net_EPA_rank, 
         Net_EPA_per_play, 
         Net_EPA_per_play_rank, 
         EPA_generated, 
         EPA_generated_rank, 
         EPA_passing, 
         EPA_passing_rank, 
         EPA_rushing, 
         EPA_rushing_rank, 
         QB_EPA, 
         QB_EPA_rank,
         redzone_EPA,
         redzone_EPA_rank,
         redzone_conversion_rate,
         redzone_conversion_rate_rank,
         sacks_allowed,
         sacks_allowed_rank,
         sack_rate,
         sack_rate_rank,
         pressures_allowed,
         pressures_allowed_rank,
         pressure_allowed_rate,
         pressure_allowed_rate_rank,
         incomplete_pass_rate,
         incomplete_pass_rate_rank,
         third_down_conversion_rate,
         third_down_conversion_rate_rank,
         EPA_passing_per_play, 
         EPA_passing_per_play_rank, 
         EPA_rushing_per_play, 
         EPA_rushing_per_play_rank, 
         QB_EPA_per_play, 
         QB_EPA_per_play_rank, 
         EPA_generated_per_play, 
         EPA_generated_per_play_rank,
         off_success_rate,
         off_success_rate_rank,
         off_rush_success_rate,
         off_rush_success_rate_rank,
         off_pass_success_rate,
         off_pass_success_rate_rank,
         off_average_air_yards_depth,
         off_average_air_yards_depth_rank,
         yards_allowed,
         yards_allowed_rank,
         passing_yards_allowed,
         passing_yards_allowed_rank,
         rushing_yards_allowed,
         rushing_yards_allowed_rank,
         EPA_given_up, 
         EPA_given_up_rank,
         EPA_passing_given,
         EPA_passing_given_rank,
         EPA_rushing_given,
         EPA_rushing_given_rank,
         QB_EPA_given,
         QB_EPA_given_rank,
         redzone_EPA_given,
         redzone_EPA_given_rank,
         redzone_conversion_rate_allowed,
         redzone_conversion_rate_allowed_rank,
         sacks_forced,
         sacks_forced_rank,
         sack_forced_rate,
         sack_forced_rate_rank,
         pressures,
         pressures_rank,
         pressure_rate,
         pressure_rate_rank,
         incomplete_pass_forced_rate,
         incomplete_pass_forced_rate_rank,
         third_down_conversion_rate_allowed,
         third_down_conversion_rate_allowed_rank,
         EPA_given_up_per_play, 
         EPA_given_up_per_play_rank,
         EPA_passing_given_per_play,
         EPA_passing_given_per_play_rank,
         EPA_rushing_given_per_play,
         EPA_rushing_given_per_play_rank,
         QB_EPA_given_per_play,
         QB_EPA_given_per_play_rank,
         def_success_rate,
         def_success_rate_rank,
         def_rush_success_rate,
         def_rush_success_rate_rank,
         def_pass_success_rate,
         def_pass_success_rate_rank,
         def_average_air_yards_depth,
         def_average_air_yards_depth_rank)
```


Chargers Data Table

```{r}
Chargers_Week_3_Summary = league_wk_3_summary %>%
  filter(team == "LAC")

DT::datatable(Chargers_Week_3_Summary)
```


Comparing Chargers Week 3 Adv Statistics to the League

```{r}
stats_to_compare = c(
  "points",
  "offensive_plays",
  "yards",
  "passing_yards",
  "rushing_yards",
  "turnover_differential",
  "offensive_turnovers",
  "Net_EPA",
  "Net_EPA_per_play",
  "EPA_generated",
  "EPA_passing",
  "EPA_rushing",
  "QB_EPA",
  "redzone_EPA",
  "redzone_conversion_rate",
  "sacks_allowed",
  "sack_rate",
  "pressures_allowed",
  "pressure_allowed_rate",
  "incomplete_pass_rate",
  "third_down_conversion_rate",
  "EPA_generated_per_play",
  "EPA_passing_per_play",
  "EPA_rushing_per_play",
  "QB_EPA_per_play",
  "off_success_rate",
  "off_rush_success_rate",
  "off_pass_success_rate",
  "off_average_air_yards_depth")

defensive_stats_to_compare = c(
  "defensive_plays",
  "defensive_takeaways",
  "yards_allowed",
  "passing_yards_allowed",
  "rushing_yards_allowed",
  "EPA_given_up",
  "EPA_passing_given",
  "EPA_rushing_given",
  "QB_EPA_given",
  "redzone_EPA_given",
  "redzone_conversion_rate_allowed",
  "sacks_forced",
  "sack_forced_rate",
  "pressures",
  "pressure_rate", 
  "incomplete_pass_forced_rate",
  "third_down_conversion_rate_allowed",
  "EPA_given_up_per_play",
  "EPA_passing_given_per_play",
  "EPA_rushing_given_per_play",
  "QB_EPA_given_per_play",
  "def_success_rate",
  "def_rush_success_rate",
  "def_pass_success_rate",
  "def_average_air_yards_depth")

Chargers_vs_League = league_wk_3_summary%>%
  filter(team == "LAC")%>%
  mutate(
    across(
      all_of(stats_to_compare),
      ~ .x - mean(league_wk_3_summary[[cur_column()]], na.rm = TRUE),
      .names = "{.col}_vs_league_avg"),
    
    across(
      all_of(defensive_stats_to_compare),
      ~ mean(league_wk_3_summary[[cur_column()]], na.rm = TRUE) - .x,
      .names = "{.col}_vs_league_avg"))
```

Data Table of Chargers week 3 stats
```{r}
DT::datatable(
  Chargers_vs_League,
  options = list(
    scrollX = TRUE,
    pageLength = 10))
```

Chargers vs League Stats Heatmap week 3:

Offense
```{r}
offense_stats = c(
  "points",
  "offensive_plays",
  "yards",
  "passing_yards",
  "rushing_yards",
  "turnover_differential",
  "offensive_turnovers",
  "Net_EPA",
  "Net_EPA_per_play",
  "EPA_generated",
  "EPA_passing",
  "EPA_rushing",
  "QB_EPA",
  "redzone_EPA",
  "redzone_conversion_rate",
  "sacks_allowed",
  "sack_rate",
  "pressures_allowed",
  "pressure_allowed_rate",
  "incomplete_pass_rate",
  "third_down_conversion_rate",
  "EPA_generated_per_play",
  "EPA_passing_per_play",
  "EPA_rushing_per_play",
  "QB_EPA_per_play",
  "off_success_rate",
  "off_rush_success_rate",
  "off_pass_success_rate",
  "off_average_air_yards_depth")

offensive_lower_is_better = c(
  "offensive_turnovers",
  "sacks_allowed",
  "sack_rate",
  "pressures_allowed",
  "pressure_allowed_rate",
  "incomplete_pass_rate")

heatmap_data = league_wk_3_summary%>%
  select(team, all_of(offense_stats))%>%
  pivot_longer(
    cols = -team,
    names_to = "stat",
    values_to = "value")%>%
  group_by(stat)%>%
  mutate(
    league_avg = mean(value, na.rm = TRUE),
    vs_league_avg = value - league_avg)%>%
  ungroup()
```

Defense
```{r}
defensive_stats = c(
  "defensive_plays",
  "defensive_takeaways",
  "yards_allowed",
  "passing_yards_allowed",
  "rushing_yards_allowed",
  "EPA_given_up",
  "EPA_passing_given",
  "EPA_rushing_given",
  "QB_EPA_given",
  "redzone_EPA_given",
  "redzone_conversion_rate_allowed",
  "sacks_forced",
  "sack_forced_rate",
  "pressures",
  "pressure_rate",
  "incomplete_pass_forced_rate",
  "third_down_conversion_rate_allowed",
  "EPA_given_up_per_play",
  "EPA_passing_given_per_play",
  "EPA_rushing_given_per_play",
  "QB_EPA_given_per_play",
  "def_success_rate",
  "def_rush_success_rate",
  "def_pass_success_rate",
  "def_average_air_yards_depth")

defensive_lower_is_better = c(
  "defensive_plays",
  "yards_allowed",
  "passing_yards_allowed",
  "rushing_yards_allowed",
  "EPA_given_up",
  "EPA_passing_given",
  "EPA_rushing_given",
  "QB_EPA_given",
  "redzone_EPA_given",
  "redzone_conversion_rate_allowed",
  "third_down_conversion_rate_allowed",
  "EPA_given_up_per_play",
  "EPA_passing_given_per_play",
  "EPA_rushing_given_per_play",
  "QB_EPA_given_per_play",
  "def_average_air_yards_depth")

heatmap_stats = c(
  offense_stats,
  defensive_stats)

heatmap_data = league_wk_3_summary%>%
  select(team, all_of(heatmap_stats))%>%
  pivot_longer(
    cols = -team,
    names_to = "stat",
    values_to = "value")%>%
  group_by(stat) %>%
  mutate(
    league_avg = mean(value, na.rm = TRUE),
    vs_league_avg = value - league_avg)%>%
  ungroup() %>%
mutate(
  performance_vs_avg = case_when(
    stat %in% defensive_lower_is_better ~ -vs_league_avg,
    stat %in% offensive_lower_is_better ~ -vs_league_avg,
    TRUE ~ vs_league_avg))%>%
  group_by(stat)%>%
  mutate(
    z_score = (
      performance_vs_avg -
        mean(performance_vs_avg, na.rm = TRUE)) /
      sd(performance_vs_avg, na.rm = TRUE))%>%
  ungroup()
```

Offensive Heatmap
```{r}
offense_heatmap_data = heatmap_data%>%
  filter(stat %in% offense_stats)

off_heatmap_wk3 = ggplot(
  offense_heatmap_data,
  aes(
    x = team,
    y = stat,
    fill = z_score))+
  
  geom_tile(color = "grey20")+
  
  scale_fill_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = 0)+
  
  labs(
    title = "NFL Week 3 Offensive Performance vs. League Average",
    subtitle = "Positive values indicate above-average performance",
    x = "Team",
    y = "Statistic",
    fill = "Vs. League Avg.")+
  
  theme_minimal()+
  theme(
    plot.background = element_rect(fill = "black", color = NA),
    panel.background = element_rect(fill = "black", color = NA),
    panel.grid = element_blank(),
    
    text = element_text(color = "white"),
    axis.text = element_text(color = "white"),
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      color = "white"),
    axis.title = element_text(color = "white"),
    
    plot.title = element_text(
      color = "white",
      face = "bold"),
    plot.subtitle = element_text(color = "grey80"),
    
    legend.background = element_rect(fill = "black", color = NA),
    legend.text = element_text(color = "white"),
    legend.title = element_text(color = "white"))

off_heatmap_wk3
```

Saving heatmap
```{r}
ggsave(
  filename = "C:/Personal Projects/Statistically Speaking/NFL_Week_3_Offensive_Heatmap.png",
  plot = off_heatmap_wk3,
  width = 14,
  height = 8,
  dpi = 300,
  bg = "black")
```


Offensive Ranking
```{r}
offensive_ranking = offense_heatmap_data%>%
  group_by(stat)%>%
  mutate(
    rank = ifelse(
      stat %in% offensive_lower_is_better,
      min_rank(value),
      min_rank(desc(value))))%>%
  ungroup()%>%
  select(team, stat, rank)%>%
  pivot_wider(
    names_from = stat,
    values_from = rank,
    names_glue = "{stat}_rank")%>%
  mutate(
    average_rank = rowMeans(
      select(., ends_with("_rank")),
      na.rm = TRUE),
    final_rank = min_rank(average_rank))%>%
  arrange(final_rank)
```

Offensive Ranking Table
```{r}
offensive_ranking_table = DT::datatable(
  offensive_ranking,
  options = list(
    scrollX = TRUE,
    pageLength = 10))

offensive_ranking_table
```

Saving table
```{r}
saveWidget(
  offensive_ranking_table,
  "C:/Personal Projects/Statistically Speaking/offensive_ranking_table_wk3.html",
  selfcontained = TRUE)
```



Defensive heatmap

```{r}

defense_heatmap_data = heatmap_data%>%
  filter(stat %in% defensive_stats)

def_heatmap_wk3 = ggplot(
  defense_heatmap_data,
  aes(
    x = team,
    y = stat,
    fill = z_score))+
  
  geom_tile(color = "grey20")+
  
  scale_fill_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = 0)+
  
  labs(
    title = "NFL Week 3 Defensive Performance vs. League Average",
    subtitle = "Positive values indicate above-average performance",
    x = "Team",
    y = "Statistic",
    fill = "Vs. League Avg.")+
  
  theme_minimal()+
  theme(
    plot.background = element_rect(fill = "black", color = NA),
    panel.background = element_rect(fill = "black", color = NA),
    panel.grid = element_blank(),
    
    text = element_text(color = "white"),
    axis.text = element_text(color = "white"),
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      color = "white"),
    axis.title = element_text(color = "white"),
    
    plot.title = element_text(
      color = "white",
      face = "bold"),
    plot.subtitle = element_text(color = "grey80"),
    
    legend.background = element_rect(fill = "black", color = NA),
    legend.text = element_text(color = "white"),
    legend.title = element_text(color = "white"))

def_heatmap_wk3
```


Saving heatmap
```{r}
ggsave(
  filename = "C:/Personal Projects/Statistically Speaking/NFL_Week_3_Defensive_Heatmap.png",
  plot = def_heatmap_wk3,
  width = 14,
  height = 8,
  dpi = 300,
  bg = "black")
```


Defensive Ranking
```{r}
defensive_ranking = defense_heatmap_data%>%
  group_by(stat)%>%
  mutate(
    rank = ifelse(
      stat %in% defensive_lower_is_better,
      min_rank(value),
      min_rank(desc(value))))%>%
  ungroup()%>%
  select(team, stat, rank)%>%
  pivot_wider(
    names_from = stat,
    values_from = rank,
    names_glue = "{stat}_rank")%>%
  mutate(
    average_rank = rowMeans(
      select(., ends_with("_rank")),
      na.rm = TRUE),
    final_rank = min_rank(average_rank))%>%
  arrange(final_rank)
```

Defensive Ranking Table
```{r}
defensive_ranking_table = DT::datatable(
  defensive_ranking,
  options = list(
    scrollX = TRUE,
    pageLength = 10))

defensive_ranking_table
```

Saving table
```{r}
saveWidget(
  defensive_ranking_table,
  "C:/Personal Projects/Statistically Speaking/defensive_ranking_table_wk3.html",
  selfcontained = TRUE)
```



Offensive Charts:
```{r}

off_rush_avg = mean(offense_wk_3$EPA_rushing, na.rm = TRUE)
off_pass_avg = mean(offense_wk_3$EPA_passing, na.rm = TRUE)

off_epa_wk_3 = ggplot(
  offense_wk_3,
  aes(
    x = EPA_rushing,
    y = EPA_passing))+
  
  nflplotR::geom_nfl_logos(
    aes(team_abbr = team),
    width = 0.075)+
  
  geom_vline(
    xintercept = off_rush_avg,
    linetype = "dashed",
    color = "grey50")+
  
  geom_hline(
    yintercept = off_pass_avg,
    linetype = "dashed",
    color = "grey50")+

  
  labs(
    title = "Team Offensive Efficiency",
    subtitle = "2026 Week 3 | Higher EPA is Better",
    x = "Rushing EPA",
    y = "Passing EPA")+
  
  theme_minimal()+
  theme(
    plot.background = element_rect(fill = "black", color = NA),
    panel.background = element_rect(fill = "black", color = NA),
    panel.grid = element_blank(),
    
    text = element_text(color = "white"),
    axis.text = element_text(color = "white"),
    axis.title = element_text(color = "white"),
    
    plot.title = element_text(
      color = "white",
      face = "bold"),
    plot.subtitle = element_text(color = "grey80"))

off_epa_wk_3
```

Saving chart
```{r}
ggsave(
  filename = "C:/Personal Projects/Statistically Speaking/NFL_Week_3_Offensive EPA Chart.png",
  plot = off_epa_wk_3,
  width = 14,
  height = 8,
  dpi = 300,
  bg = "black")
```


Success Rate

```{r}

off_rush_suc_avg = mean(offense_wk_3$off_rush_success_rate, na.rm = TRUE)
off_pass_suc_avg = mean(offense_wk_3$off_pass_success_rate, na.rm = TRUE)

off_suc_wk_3 = ggplot(
  offense_wk_3,
  aes(
    x = off_rush_success_rate,
    y = off_pass_success_rate))+
  
  nflplotR::geom_nfl_logos(
    aes(team_abbr = team),
    width = 0.075)+
  
  geom_vline(
    xintercept = off_rush_suc_avg,
    linetype = "dashed",
    color = "grey50")+
  
  geom_hline(
    yintercept = off_pass_suc_avg,
    linetype = "dashed",
    color = "grey50")+

  
  labs(
    title = "Team Offensive Success",
    subtitle = "2026 Week 3 | Higher Success Rate is Better",
    x = "Rushing Success %",
    y = "Passing Success %")+
  
  theme_minimal()+
  theme(
    plot.background = element_rect(fill = "black", color = NA),
    panel.background = element_rect(fill = "black", color = NA),
    panel.grid = element_blank(),
    
    text = element_text(color = "white"),
    axis.text = element_text(color = "white"),
    axis.title = element_text(color = "white"),
    
    plot.title = element_text(
      color = "white",
      face = "bold"),
    plot.subtitle = element_text(color = "grey80"))

off_suc_wk_3
```

Saving chart
```{r}
ggsave(
  filename = "C:/Personal Projects/Statistically Speaking/NFL_Week_3_Offensive Success Rate Chart.png",
  plot = off_suc_wk_3,
  width = 14,
  height = 8,
  dpi = 300,
  bg = "black")
```



Defensive Charts:

```{r}

def_rush_avg = mean(defense_wk_3$EPA_rushing_given, na.rm = TRUE)
def_pass_avg = mean(defense_wk_3$EPA_passing_given, na.rm = TRUE)

def_epa_wk_3 = ggplot(
  defense_wk_3,
  aes(
    x = EPA_rushing_given,
    y = EPA_passing_given))+
  
  nflplotR::geom_nfl_logos(
    aes(team_abbr = team),
    width = 0.075)+
  
  geom_vline(
    xintercept = def_rush_avg,
    linetype = "dashed",
    color = "grey50")+
  
  geom_hline(
    yintercept = def_pass_avg,
    linetype = "dashed",
    color = "grey50")+
  
  scale_x_reverse()+
  scale_y_reverse()+
  
  labs(
    title = "Team Defensive Efficiency",
    subtitle = "2026 Week 3 | Lower EPA Allowed is Better",
    x = "Rushing EPA Allowed",
    y = "Passing EPA Allowed")+
  
  theme_minimal()+
  theme(
    plot.background = element_rect(fill = "black", color = NA),
    panel.background = element_rect(fill = "black", color = NA),
    panel.grid = element_blank(),
    
    text = element_text(color = "white"),
    axis.text = element_text(color = "white"),
    axis.title = element_text(color = "white"),
    
    plot.title = element_text(
      color = "white",
      face = "bold"),
    plot.subtitle = element_text(color = "grey80"))

def_epa_wk_3
```

Saving chart
```{r}
ggsave(
  filename = "C:/Personal Projects/Statistically Speaking/NFL_Week_3_Defensive EPA Chart.png",
  plot = def_epa_wk_3,
  width = 14,
  height = 8,
  dpi = 300,
  bg = "black")
```



Defensive Success Rate

Success Rate

```{r}

def_rush_suc_avg = mean(defense_wk_3$def_rush_success_rate, na.rm = TRUE)
def_pass_suc_avg = mean(defense_wk_3$def_pass_success_rate, na.rm = TRUE)

def_suc_wk_3 = ggplot(
  defense_wk_3,
  aes(
    x = def_rush_success_rate,
    y = def_pass_success_rate))+
  
  nflplotR::geom_nfl_logos(
    aes(team_abbr = team),
    width = 0.075)+
  
  geom_vline(
    xintercept = def_rush_suc_avg,
    linetype = "dashed",
    color = "grey50")+
  
  geom_hline(
    yintercept = def_pass_suc_avg,
    linetype = "dashed",
    color = "grey50")+

  
  labs(
    title = "Team Defensive Success",
    subtitle = "2026 Week 3 | Higher Success Rate is Better",
    x = "Rushing Success %",
    y = "Passing Success %")+
  
  theme_minimal()+
  theme(
    plot.background = element_rect(fill = "black", color = NA),
    panel.background = element_rect(fill = "black", color = NA),
    panel.grid = element_blank(),
    
    text = element_text(color = "white"),
    axis.text = element_text(color = "white"),
    axis.title = element_text(color = "white"),
    
    plot.title = element_text(
      color = "white",
      face = "bold"),
    plot.subtitle = element_text(color = "grey80"))

def_suc_wk_3
```

Saving chart
```{r}
ggsave(
  filename = "C:/Personal Projects/Statistically Speaking/NFL_Week_3_Defensive Success Rate Chart.png",
  plot = def_suc_wk_3,
  width = 14,
  height = 8,
  dpi = 300,
  bg = "black")
```





Comparing week 2 to week 3:

```{r}
league_wk_2 = league%>%
  filter(week == 2)

team_points_wk_2 = league_wk_2%>%
  group_by(game_id)%>%
  summarise(
    home_team = first(home_team),
    away_team = first(away_team),
    home_points = max(total_home_score, na.rm = TRUE),
    away_points = max(total_away_score, na.rm = TRUE),
    .groups = "drop")

team_points_wk_2 = bind_rows(
  team_points_wk_2 %>%
    transmute(
      team = home_team,
      points = home_points),
  
  team_points_wk_2 %>%
    transmute(
      team = away_team,
      points = away_points))
#Offense
offense_wk_2 = league_wk_2%>%
  filter(
    !is.na(posteam),
    !play_type %in% c(
      "kickoff",
      "extra_point",
      "field_goal"))%>%
  group_by(posteam)%>%
  summarise(
    
    yards = sum(yards_gained, na.rm = TRUE),
    
    passing_yards = sum(passing_yards, na.rm = TRUE),
    
    rushing_yards = sum(rushing_yards, na.rm = TRUE),
    
    offensive_turnovers = sum(
      interception + fumble_lost,
      na.rm = TRUE),
    
    EPA_generated = sum(epa, na.rm = TRUE),
    
    EPA_passing = sum(
      epa[play_type == "pass"],
      na.rm = TRUE),
    
    EPA_rushing = sum(
      epa[play_type == "run"],
      na.rm = TRUE),
    
    QB_EPA = sum(qb_epa, na.rm = TRUE),
    
    redzone_EPA = sum(
      epa[redzone == 1],
      na.rm = TRUE),
    
    redzone_drives = n_distinct(
      drive[redzone == 1]),
    
    redzone_touchdowns = n_distinct(
      drive[redzone == 1 & touchdown == 1]),
    
    redzone_conversion_rate =
      redzone_touchdowns / redzone_drives,
    
    sacks_allowed = sum(sack, na.rm = TRUE),
    
    pressures_allowed = sum(qb_hit, na.rm = TRUE),
    
    incomplete_passes = sum(incomplete_pass, na.rm = TRUE),
    
    offensive_plays = n(),
    
    pass_plays = sum(
      play_type == "pass",
      na.rm = TRUE),
    
    rush_plays = sum(
      play_type == "run",
      na.rm = TRUE),
    
    sack_rate =
      sacks_allowed / pass_plays,
    
    pressure_allowed_rate = pressures_allowed / pass_plays, 
    
    incomplete_pass_rate =
      incomplete_passes / pass_plays,
    
    third_downs_faced =
      sum(down == 3, na.rm = TRUE),
    
    third_down_conversion_rate =
      sum(third_down_converted, na.rm = TRUE) /
      third_downs_faced,
    
    EPA_generated_per_play =
      EPA_generated / offensive_plays,
    
    EPA_passing_per_play =
      EPA_passing / pass_plays,
    
    EPA_rushing_per_play =
      EPA_rushing / rush_plays,
    
    QB_EPA_per_play =
      QB_EPA / pass_plays,
    
    off_success_rate =
      sum(success, na.rm = TRUE) /
      offensive_plays,
    
    off_rush_success_rate =
      sum(success[play_type == "run"], na.rm = TRUE) /
      rush_plays,
    
    off_pass_success_rate =
      sum(success[play_type == "pass"], na.rm = TRUE) /
      pass_plays,
    
    off_average_air_yards_depth =
      mean(air_yards, na.rm = TRUE),
    
    .groups = "drop")%>%
  rename(team = posteam)

#Defense
defense_wk_2 = league_wk_2%>%
  filter(
    !is.na(defteam),
    !play_type %in% c(
      "kickoff",
      "extra_point",
      "field_goal"))%>%
  group_by(defteam)%>%
  summarise(
    
    yards_allowed = sum(
      yards_gained,
      na.rm = TRUE),
    
    passing_yards_allowed = sum(
      passing_yards,
      na.rm = TRUE),
    
    rushing_yards_allowed = sum(
      rushing_yards,
      na.rm = TRUE),
    
    defensive_takeaways = sum(
      interception + fumble_forced,
      na.rm = TRUE),
    
    EPA_given_up = sum(
      epa,
      na.rm = TRUE),
    
    EPA_passing_given = sum(
      epa[play_type == "pass"],
      na.rm = TRUE),
    
    EPA_rushing_given = sum(
      epa[play_type == "run"],
      na.rm = TRUE),
    
    QB_EPA_given = sum(
      qb_epa,
      na.rm = TRUE),
    
    redzone_EPA_given = sum(
      epa[redzone == 1],
      na.rm = TRUE),
    
    redzone_drives_allowed = n_distinct(
      drive[redzone == 1]),
    
    redzone_touchdowns_allowed = n_distinct(
      drive[redzone == 1 & touchdown == 1]),
    
    redzone_conversion_rate_allowed =
      redzone_touchdowns_allowed /
      redzone_drives_allowed,
    
    sacks_forced = sum(
      sack,
      na.rm = TRUE),
    
    pressures = sum(qb_hit, na.rm = TRUE),
    
    incomplete_passes_forced = sum(
      incomplete_pass,
      na.rm = TRUE),
    
    defensive_plays = n(),
    
    pass_plays_faced = sum(
      play_type == "pass",
      na.rm = TRUE),
    
    rush_plays_faced = sum(
      play_type == "run",
      na.rm = TRUE),
    
    sack_forced_rate =
      sacks_forced / pass_plays_faced,
    
    pressure_rate = pressures / pass_plays_faced,
    
    incomplete_pass_forced_rate =
      incomplete_passes_forced /
      pass_plays_faced,
    
    third_downs_forced =
      sum(down == 3, na.rm = TRUE),
    
    third_down_conversion_rate_allowed =
      sum(third_down_converted, na.rm = TRUE) /
      third_downs_forced,
    
    EPA_given_up_per_play =
      EPA_given_up / defensive_plays,
    
    EPA_passing_given_per_play =
      EPA_passing_given / pass_plays_faced,
    
    EPA_rushing_given_per_play =
      EPA_rushing_given / rush_plays_faced,
    
    QB_EPA_given_per_play =
      QB_EPA_given / pass_plays_faced,
    
    def_success_rate =
      1 - (
        sum(success, na.rm = TRUE) /
        defensive_plays),
    
    def_rush_success_rate =
      1 - (
        sum(success[play_type == "run"], na.rm = TRUE) /
        rush_plays_faced),
    
    def_pass_success_rate =
      1 - (
        sum(success[play_type == "pass"], na.rm = TRUE) /
        pass_plays_faced),
    
    def_average_air_yards_depth =
      mean(air_yards, na.rm = TRUE),
    
    .groups = "drop")%>%
  rename(team = defteam)


```

Joined Data
```{r}
league_wk_2_summary = offense_wk_2%>%
  left_join(
    defense_wk_2,
    by = "team")%>%
  left_join(
    team_points_wk_2,
    by = "team")%>%
  mutate(
    Net_EPA =
      EPA_generated - EPA_given_up,
    
    Net_EPA_per_play =
      EPA_generated_per_play -
      EPA_given_up_per_play,
    
    turnover_differential =
      defensive_takeaways - offensive_turnovers)%>%
  mutate(
    points_rank = min_rank(desc(points)),
    
    offensive_plays_rank = min_rank(desc(offensive_plays)),
    
    defensive_plays_rank = min_rank((defensive_plays)),
    
    yards_rank = min_rank(desc(yards)),
    
    passing_yards_rank = min_rank(desc(passing_yards)),
    
    rushing_yards_rank = min_rank(desc(rushing_yards)),
    
    offensive_turnovers_rank = min_rank(offensive_turnovers),
    
    EPA_generated_rank = min_rank(desc(EPA_generated)),
    
    EPA_passing_rank = min_rank(desc(EPA_passing)),
    
    EPA_rushing_rank = min_rank(desc(EPA_rushing)),
    
    QB_EPA_rank = min_rank(desc(QB_EPA)),
    
    redzone_EPA_rank = min_rank(desc(redzone_EPA)),
    
    redzone_conversion_rate_rank = min_rank(desc(redzone_conversion_rate)),
    
    sacks_allowed_rank = min_rank(sacks_allowed),
    
    sack_rate_rank = min_rank(sack_rate),
    
    pressures_allowed_rank = min_rank(pressures_allowed),
    
    pressure_allowed_rate_rank = min_rank(pressure_allowed_rate),
    
    incomplete_pass_rate_rank = min_rank(incomplete_pass_rate),
    
    third_down_conversion_rate_rank = min_rank(desc(third_down_conversion_rate)),
    
    EPA_generated_per_play_rank = min_rank(desc(EPA_generated_per_play)),
    
    EPA_passing_per_play_rank = min_rank(desc(EPA_passing_per_play)),
    
    EPA_rushing_per_play_rank = min_rank(desc(EPA_rushing_per_play)),
    
    QB_EPA_per_play_rank = min_rank(desc(QB_EPA_per_play)),
    
    off_success_rate_rank = min_rank(desc(off_success_rate)),
    
    off_rush_success_rate_rank = min_rank(desc(off_rush_success_rate)),
    
    off_pass_success_rate_rank = min_rank(desc(off_pass_success_rate)),
    
    off_average_air_yards_depth_rank = min_rank(desc(off_average_air_yards_depth)),
    
    yards_allowed_rank = min_rank(yards_allowed),
    
    passing_yards_allowed_rank = min_rank((passing_yards_allowed)),
    
    rushing_yards_allowed_rank = min_rank((rushing_yards_allowed)),
    
    defensive_takeaways_rank = min_rank(desc(defensive_takeaways)),
    
    EPA_given_up_rank = min_rank(EPA_given_up),
    
    EPA_passing_given_rank = min_rank(EPA_passing_given),
    
    EPA_rushing_given_rank = min_rank(EPA_rushing_given),
    
    QB_EPA_given_rank = min_rank(QB_EPA_given),
    
    redzone_EPA_given_rank = min_rank(redzone_EPA_given),
    
    redzone_conversion_rate_allowed_rank = min_rank(redzone_conversion_rate_allowed),
    
    sacks_forced_rank = min_rank(desc(sacks_forced)),
    
    sack_forced_rate_rank = min_rank(desc(sack_forced_rate)),
    
    pressures_rank = min_rank(desc(pressures)),
    
    pressure_rate_rank = min_rank(desc(pressure_rate)),
    
    incomplete_pass_forced_rate_rank = min_rank(desc(incomplete_pass_forced_rate)),
    
    third_down_conversion_rate_allowed_rank = min_rank(desc(third_down_conversion_rate_allowed)),
    
    EPA_given_up_per_play_rank = min_rank(EPA_given_up_per_play),
    
    EPA_passing_given_per_play_rank = min_rank(EPA_passing_given_per_play),
     
    EPA_rushing_given_per_play_rank = min_rank(EPA_rushing_given_per_play),
    
    QB_EPA_given_per_play_rank = min_rank(QB_EPA_given_per_play),
    
    def_success_rate_rank = min_rank(desc(def_success_rate)),
    
    def_rush_success_rate_rank = min_rank(desc(def_rush_success_rate)),
    
    def_pass_success_rate_rank = min_rank(desc(def_pass_success_rate)),
    
    def_average_air_yards_depth_rank = min_rank((def_average_air_yards_depth)),
    
    Net_EPA_rank = min_rank(desc(Net_EPA)),
    
    Net_EPA_per_play_rank = min_rank(desc(Net_EPA_per_play)),
    
    turnover_differential_rank = min_rank(desc(turnover_differential)))


```


Cleaning it up

```{r}
league_wk_2_summary = league_wk_2_summary%>%
  select(team, 
         points, 
         points_rank, 
         offensive_plays, 
         offensive_plays_rank, 
         yards,
         yards_rank, 
         passing_yards, 
         passing_yards_rank, 
         rushing_yards, 
         rushing_yards_rank, 
         turnover_differential, 
         turnover_differential_rank, 
         offensive_turnovers, 
         offensive_turnovers_rank, 
         defensive_plays, 
         defensive_takeaways, 
         defensive_takeaways_rank, 
         Net_EPA, 
         Net_EPA_rank, 
         Net_EPA_per_play, 
         Net_EPA_per_play_rank, 
         EPA_generated, 
         EPA_generated_rank, 
         EPA_passing, 
         EPA_passing_rank, 
         EPA_rushing, 
         EPA_rushing_rank, 
         QB_EPA, 
         QB_EPA_rank,
         redzone_EPA,
         redzone_EPA_rank,
         redzone_conversion_rate,
         redzone_conversion_rate_rank,
         sacks_allowed,
         sacks_allowed_rank,
         sack_rate,
         sack_rate_rank,
         pressures_allowed,
         pressures_allowed_rank,
         pressure_allowed_rate,
         pressure_allowed_rate_rank,
         incomplete_pass_rate,
         incomplete_pass_rate_rank,
         third_down_conversion_rate,
         third_down_conversion_rate_rank,
         EPA_passing_per_play, 
         EPA_passing_per_play_rank, 
         EPA_rushing_per_play, 
         EPA_rushing_per_play_rank, 
         QB_EPA_per_play, 
         QB_EPA_per_play_rank, 
         EPA_generated_per_play, 
         EPA_generated_per_play_rank,
         off_success_rate,
         off_success_rate_rank,
         off_rush_success_rate,
         off_rush_success_rate_rank,
         off_pass_success_rate,
         off_pass_success_rate_rank,
         off_average_air_yards_depth,
         off_average_air_yards_depth_rank,
         yards_allowed,
         yards_allowed_rank,
         passing_yards_allowed,
         passing_yards_allowed_rank,
         rushing_yards_allowed,
         rushing_yards_allowed_rank,
         EPA_given_up, 
         EPA_given_up_rank,
         EPA_passing_given,
         EPA_passing_given_rank,
         EPA_rushing_given,
         EPA_rushing_given_rank,
         QB_EPA_given,
         QB_EPA_given_rank,
         redzone_EPA_given,
         redzone_EPA_given_rank,
         redzone_conversion_rate_allowed,
         redzone_conversion_rate_allowed_rank,
         sacks_forced,
         sacks_forced_rank,
         sack_forced_rate,
         sack_forced_rate_rank,
         pressures,
         pressures_rank,
         pressure_rate,
         pressure_rate_rank, 
         incomplete_pass_forced_rate,
         incomplete_pass_forced_rate_rank,
         third_down_conversion_rate_allowed,
         third_down_conversion_rate_allowed_rank,
         EPA_given_up_per_play, 
         EPA_given_up_per_play_rank,
         EPA_passing_given_per_play,
         EPA_passing_given_per_play_rank,
         EPA_rushing_given_per_play,
         EPA_rushing_given_per_play_rank,
         QB_EPA_given_per_play,
         QB_EPA_given_per_play_rank,
         def_success_rate,
         def_success_rate_rank,
         def_rush_success_rate,
         def_rush_success_rate_rank,
         def_pass_success_rate,
         def_pass_success_rate_rank,
         def_average_air_yards_depth,
         def_average_air_yards_depth_rank)
```

Joining with Week 2 Summary

```{r}
offensive_stats = c(
  "points",
  "offensive_plays",
  "yards",
  "passing_yards",
  "rushing_yards",
  "turnover_differential",
  "offensive_turnovers",
  "Net_EPA",
  "Net_EPA_per_play",
  "EPA_generated",
  "EPA_passing",
  "EPA_rushing",
  "QB_EPA",
  "redzone_EPA",
  "redzone_conversion_rate",
  "sacks_allowed",
  "sack_rate",
  "pressures_allowed",
  "pressure_allowed_rate",
  "incomplete_pass_rate",
  "third_down_conversion_rate",
  "EPA_generated_per_play",
  "EPA_passing_per_play",
  "EPA_rushing_per_play",
  "QB_EPA_per_play",
  "off_success_rate",
  "off_rush_success_rate",
  "off_pass_success_rate",
  "off_average_air_yards_depth")

defensive_stats = c(
  "defensive_plays",
  "defensive_takeaways",
  "yards_allowed",
  "passing_yards_allowed",
  "rushing_yards_allowed",
  "EPA_given_up",
  "EPA_passing_given",
  "EPA_rushing_given",
  "QB_EPA_given",
  "redzone_EPA_given",
  "redzone_conversion_rate_allowed",
  "sacks_forced",
  "sack_forced_rate",
  "pressures",
  "pressure_rate", 
  "incomplete_pass_forced_rate",
  "third_down_conversion_rate_allowed",
  "EPA_given_up_per_play",
  "EPA_passing_given_per_play",
  "EPA_rushing_given_per_play",
  "QB_EPA_given_per_play",
  "def_success_rate",
  "def_rush_success_rate",
  "def_pass_success_rate",
  "def_average_air_yards_depth")


Chargers_wk2_vs_wk3 = league_wk_2_summary%>%
  filter(team == "LAC")%>%
  select(
    team,
    all_of(c(offensive_stats, defensive_stats)))%>%
  rename_with(
    ~ paste0(.x, "_wk2"),
    -team)%>%
  left_join(
    league_wk_3_summary%>%
      filter(team == "LAC")%>%
      select(
        team,
        all_of(c(offensive_stats, defensive_stats)))%>%
      rename_with(
        ~ paste0(.x, "_wk3"),
        -team),
    by = "team")%>%
mutate(
  # Higher is better
  points_change_from_week_2 =
    points_wk3 - points_wk2,

  offensive_plays_change_from_week_2 =
    offensive_plays_wk3 - offensive_plays_wk2,

  yards_change_from_week_2 =
    yards_wk3 - yards_wk2,

  passing_yards_change_from_week_2 =
    passing_yards_wk3 - passing_yards_wk2,

  rushing_yards_change_from_week_2 =
    rushing_yards_wk3 - rushing_yards_wk2,

  turnover_differential_change_from_week_2 =
    turnover_differential_wk3 - turnover_differential_wk2,

  defensive_takeaways_change_from_week_2 =
    defensive_takeaways_wk3 - defensive_takeaways_wk2,

  Net_EPA_change_from_week_2 =
    Net_EPA_wk3 - Net_EPA_wk2,

  Net_EPA_per_play_change_from_week_2 =
    Net_EPA_per_play_wk3 - Net_EPA_per_play_wk2,

  EPA_generated_change_from_week_2 =
    EPA_generated_wk3 - EPA_generated_wk2,

  EPA_passing_change_from_week_2 =
    EPA_passing_wk3 - EPA_passing_wk2,

  EPA_rushing_change_from_week_2 =
    EPA_rushing_wk3 - EPA_rushing_wk2,

  QB_EPA_change_from_week_2 =
    QB_EPA_wk3 - QB_EPA_wk2,
  
  redzone_EPA_change_from_week_2 =
    redzone_EPA_wk3 - redzone_EPA_wk2,
  
  redzone_conversion_rate_change_from_week_2 =
    redzone_conversion_rate_wk3 - redzone_conversion_rate_wk2,
  
  third_down_conversion_rate_change_from_week_2 =
    third_down_conversion_rate_wk3 - third_down_conversion_rate_wk2,
  
  EPA_generated_per_play_change_from_week_2 =
    EPA_generated_per_play_wk3 - EPA_generated_per_play_wk2,

  EPA_passing_per_play_change_from_week_2 =
    EPA_passing_per_play_wk3 - EPA_passing_per_play_wk2,

  EPA_rushing_per_play_change_from_week_2 =
    EPA_rushing_per_play_wk3 - EPA_rushing_per_play_wk2,

  QB_EPA_per_play_change_from_week_2 =
    QB_EPA_per_play_wk3 - QB_EPA_per_play_wk2,
  
  off_success_rate_change_from_week_2 =
    off_success_rate_wk3 - off_success_rate_wk2,
  
  off_rush_success_rate_change_from_week_2 =
    off_rush_success_rate_wk3 - off_rush_success_rate_wk2,
  
  off_pass_success_rate_change_from_week_2 =
    off_pass_success_rate_wk3 - off_pass_success_rate_wk2,
  
  def_success_rate_change_from_week_2 =
    def_success_rate_wk3 - def_success_rate_wk2,
  
  def_rush_success_rate_change_from_week_2 =
    def_rush_success_rate_wk3 - def_rush_success_rate_wk2,
  
  def_pass_success_rate_change_from_week_2 =
    def_pass_success_rate_wk3 - def_pass_success_rate_wk2,
  
  off_average_air_yards_depth_change_from_week_2 =
    off_average_air_yards_depth_wk3 - off_average_air_yards_depth_wk2,
  
  sacks_forced_change_from_week_2 =
    sacks_forced_wk3 - sacks_forced_wk2,
  
  sack_forced_rate_change_from_week_2 =
    sack_forced_rate_wk3 - sack_forced_rate_wk2,
  
  pressures_change_from_week_2 =
    pressures_wk3 - pressures_wk2,
  
  pressure_rate_change_from_week_2 =
    pressure_rate_wk3 - pressure_rate_wk2,
  
  incomplete_pass_forced_rate_change_from_week_2 =
    incomplete_pass_forced_rate_wk3 - incomplete_pass_forced_rate_wk2,
  

  # Lower is better — reverse the subtraction
  defensive_plays_change_from_week_2 =
    defensive_plays_wk3 - defensive_plays_wk2, 
  
  sacks_allowed_change_from_week_2 =
    sacks_allowed_wk2 - sacks_allowed_wk3,
  
  sack_rate_change_from_week_2 =
    sack_rate_wk2 - sack_rate_wk3,
  
  pressures_allowed_change_from_week_2 =
    pressures_allowed_wk2 - pressures_allowed_wk3,
  
  pressure_allowed_rate_change_from_week_2 =
    pressure_allowed_rate_wk2 - pressure_allowed_rate_wk3,
  
  incomplete_pass_rate_change_from_week_2 =
    incomplete_pass_rate_wk2 - incomplete_pass_rate_wk3,
  
  yards_allowed_change_from_week_2 =
    yards_allowed_wk2 - yards_allowed_wk3,

  passing_yards_allowed_change_from_week_2 =
    passing_yards_allowed_wk2 - passing_yards_allowed_wk3,

  rushing_yards_allowed_change_from_week_2 =
    rushing_yards_allowed_wk2 - rushing_yards_allowed_wk3,

  offensive_turnovers_change_from_week_2 =
    offensive_turnovers_wk2 - offensive_turnovers_wk3,

  EPA_given_up_change_from_week_2 =
    EPA_given_up_wk2 - EPA_given_up_wk3,

  EPA_passing_given_change_from_week_2 =
    EPA_passing_given_wk2 - EPA_passing_given_wk3,

  EPA_rushing_given_change_from_week_2 =
    EPA_rushing_given_wk2 - EPA_rushing_given_wk3,

  QB_EPA_given_change_from_week_2 =
    QB_EPA_given_wk2 - QB_EPA_given_wk3,
  
  redzone_EPA_given_change_from_week_2 = 
    redzone_EPA_given_wk2 - redzone_EPA_given_wk3,
  
  redzone_conversion_rate_allowed_change_from_week_2 =
    redzone_conversion_rate_allowed_wk2 - redzone_conversion_rate_allowed_wk3,
  
  third_down_conversion_rate_allowed_change_from_wk2 =
    third_down_conversion_rate_allowed_wk2 - third_down_conversion_rate_allowed_wk3,

  EPA_given_up_per_play_change_from_week_2 =
    EPA_given_up_per_play_wk2 - EPA_given_up_per_play_wk3,

  EPA_passing_given_per_play_change_from_week_2 =
    EPA_passing_given_per_play_wk2 - EPA_passing_given_per_play_wk3,

  EPA_rushing_given_per_play_change_from_week_2 =
    EPA_rushing_given_per_play_wk2 - EPA_rushing_given_per_play_wk3,

  QB_EPA_given_per_play_change_from_week_2 =
    QB_EPA_given_per_play_wk2 - QB_EPA_given_per_play_wk3,
  
  def_average_air_yards_depth_change_from_week_2 =
    def_average_air_yards_depth_wk3 - def_average_air_yards_depth_wk2)
```


#Week 1 vs Week 2 Data Table:
```{r}
DT::datatable(
  Chargers_wk2_vs_wk3,
  options = list(
    scrollX = TRUE,
    pageLength = 10))
```

Week 3 Week 2 Difference HeatMap Data

```{r}
Chargers_wk2_wk3_table = tibble(
  
  Stat = c(
    # Offense
    "Points",
    "Yards",
    "Passing Yards",
    "Rushing Yards",
    "Offensive Turnovers",
    "Turnover Differential",
    "Net EPA",
    "Net EPA / Play",
    "EPA Generated",
    "EPA Passing",
    "EPA Rushing",
    "QB EPA",
    "redzone EPA",
    "redzone conversion rate",
    "sacks allowed",
    "sack rate",
    "pressures allowed",
    "pressure allowed rate",
    "incomplete pass rate",
    "third down conversion rate",
    "EPA Generated / Play",
    "EPA Passing / Play",
    "EPA Rushing / Play",
    "QB EPA / Play",
    "Off Success Rate",
    "Off Rush Success Rate",
    "Off Pass Success Rate",
    "Off Average Air Yards Depth",
    
    # Defense
    "Defensive Plays",
    "Yards Allowed",
    "Passing Yards Allowed",
    "Rushing Yards Allowed",
    "Defensive Turnovers",
    "EPA Given Up",
    "EPA Passing Given",
    "EPA Rushing Given",
    "QB EPA Given",
    "redzone EPA Given",
    "redzone conversion rate allowed",
    "sacks forced",
    "sack forced rate",
    "pressures",
    "pressure rate",
    "incomplete pass forced rate",
    "third down conversion rate allowed",
    "EPA Given Up / Play",
    "EPA Passing Given / Play",
    "EPA Rushing Given / Play",
    "QB EPA Given / Play",
    "Def Success Rate",
    "Def Rush Success Rate",
    "Def Pass Success Rate",
    "Def Average Air Yards Depth"),
  
  week_2 = c(
    # Offense
    Chargers_wk2_vs_wk3$points_wk2,
    Chargers_wk2_vs_wk3$yards_wk2,
    Chargers_wk2_vs_wk3$passing_yards_wk2,
    Chargers_wk2_vs_wk3$rushing_yards_wk2,
    Chargers_wk2_vs_wk3$offensive_turnovers_wk2,
    Chargers_wk2_vs_wk3$turnover_differential_wk2,
    Chargers_wk2_vs_wk3$Net_EPA_wk2,
    Chargers_wk2_vs_wk3$Net_EPA_per_play_wk2,
    Chargers_wk2_vs_wk3$EPA_generated_wk2,
    Chargers_wk2_vs_wk3$EPA_passing_wk2,
    Chargers_wk2_vs_wk3$EPA_rushing_wk2,
    Chargers_wk2_vs_wk3$QB_EPA_wk2,
    Chargers_wk2_vs_wk3$redzone_EPA_wk2,
    Chargers_wk2_vs_wk3$redzone_conversion_rate_wk2,
    Chargers_wk2_vs_wk3$sacks_allowed_wk2,
    Chargers_wk2_vs_wk3$sack_rate_wk2,
    Chargers_wk2_vs_wk3$pressures_allowed_wk2,
    Chargers_wk2_vs_wk3$pressure_allowed_rate_wk2,
    Chargers_wk2_vs_wk3$incomplete_pass_rate_wk2,
    Chargers_wk2_vs_wk3$third_down_conversion_rate_wk2,
    Chargers_wk2_vs_wk3$EPA_generated_per_play_wk2,
    Chargers_wk2_vs_wk3$EPA_passing_per_play_wk2,
    Chargers_wk2_vs_wk3$EPA_rushing_per_play_wk2,
    Chargers_wk2_vs_wk3$QB_EPA_per_play_wk2,
    Chargers_wk2_vs_wk3$off_success_rate_wk2,
    Chargers_wk2_vs_wk3$off_rush_success_rate_wk2,
    Chargers_wk2_vs_wk3$off_pass_success_rate_wk2,
    Chargers_wk2_vs_wk3$off_average_air_yards_depth_wk2,
    
    # Defense
    Chargers_wk2_vs_wk3$defensive_plays_wk2,
    Chargers_wk2_vs_wk3$yards_allowed_wk2,
    Chargers_wk2_vs_wk3$passing_yards_allowed_wk2,
    Chargers_wk2_vs_wk3$rushing_yards_allowed_wk2,
    Chargers_wk2_vs_wk3$defensive_takeaways_wk2,
    Chargers_wk2_vs_wk3$EPA_given_up_wk2,
    Chargers_wk2_vs_wk3$EPA_passing_given_wk2,
    Chargers_wk2_vs_wk3$EPA_rushing_given_wk2,
    Chargers_wk2_vs_wk3$QB_EPA_given_wk2,
    Chargers_wk2_vs_wk3$redzone_EPA_given_wk2,
    Chargers_wk2_vs_wk3$redzone_conversion_rate_allowed_wk2,
    Chargers_wk2_vs_wk3$sacks_forced_wk2,
    Chargers_wk2_vs_wk3$sack_forced_rate_wk2,
    Chargers_wk2_vs_wk3$pressures_wk2,
    Chargers_wk2_vs_wk3$pressure_rate_wk2,
    Chargers_wk2_vs_wk3$incomplete_pass_forced_rate_wk2,
    Chargers_wk2_vs_wk3$third_down_conversion_rate_allowed_wk2,
    Chargers_wk2_vs_wk3$EPA_given_up_per_play_wk2,
    Chargers_wk2_vs_wk3$EPA_passing_given_per_play_wk2,
    Chargers_wk2_vs_wk3$EPA_rushing_given_per_play_wk2,
    Chargers_wk2_vs_wk3$QB_EPA_given_per_play_wk2,
    Chargers_wk2_vs_wk3$def_success_rate_wk2,
    Chargers_wk2_vs_wk3$def_rush_success_rate_wk2,
    Chargers_wk2_vs_wk3$def_pass_success_rate_wk2,
    Chargers_wk2_vs_wk3$def_average_air_yards_depth_wk2),
  
  Week_3 = c(
    # Offense
    Chargers_wk2_vs_wk3$points_wk3,
    Chargers_wk2_vs_wk3$yards_wk3,
    Chargers_wk2_vs_wk3$passing_yards_wk3,
    Chargers_wk2_vs_wk3$rushing_yards_wk3,
    Chargers_wk2_vs_wk3$offensive_turnovers_wk3,
    Chargers_wk2_vs_wk3$turnover_differential_wk3,
    Chargers_wk2_vs_wk3$Net_EPA_wk3,
    Chargers_wk2_vs_wk3$Net_EPA_per_play_wk3,
    Chargers_wk2_vs_wk3$EPA_generated_wk3,
    Chargers_wk2_vs_wk3$EPA_passing_wk3,
    Chargers_wk2_vs_wk3$EPA_rushing_wk3,
    Chargers_wk2_vs_wk3$QB_EPA_wk3,
    Chargers_wk2_vs_wk3$redzone_EPA_wk3,
    Chargers_wk2_vs_wk3$redzone_conversion_rate_wk3,
    Chargers_wk2_vs_wk3$sacks_allowed_wk3,
    Chargers_wk2_vs_wk3$sack_rate_wk3,
    Chargers_wk2_vs_wk3$pressures_allowed_wk3,
    Chargers_wk2_vs_wk3$pressure_allowed_rate_wk3,
    Chargers_wk2_vs_wk3$incomplete_pass_rate_wk3,
    Chargers_wk2_vs_wk3$third_down_conversion_rate_wk3,
    Chargers_wk2_vs_wk3$EPA_generated_per_play_wk3,
    Chargers_wk2_vs_wk3$EPA_passing_per_play_wk3,
    Chargers_wk2_vs_wk3$EPA_rushing_per_play_wk3,
    Chargers_wk2_vs_wk3$QB_EPA_per_play_wk3,
    Chargers_wk2_vs_wk3$off_success_rate_wk3,
    Chargers_wk2_vs_wk3$off_rush_success_rate_wk3,
    Chargers_wk2_vs_wk3$off_pass_success_rate_wk3,
    Chargers_wk2_vs_wk3$off_average_air_yards_depth_wk3,
    
    # Defense
    Chargers_wk2_vs_wk3$defensive_plays_wk3,
    Chargers_wk2_vs_wk3$yards_allowed_wk3,
    Chargers_wk2_vs_wk3$passing_yards_allowed_wk3,
    Chargers_wk2_vs_wk3$rushing_yards_allowed_wk3,
    Chargers_wk2_vs_wk3$defensive_takeaways_wk3,
    Chargers_wk2_vs_wk3$EPA_given_up_wk3,
    Chargers_wk2_vs_wk3$EPA_passing_given_wk3,
    Chargers_wk2_vs_wk3$EPA_rushing_given_wk3,
    Chargers_wk2_vs_wk3$QB_EPA_given_wk3,
    Chargers_wk2_vs_wk3$redzone_EPA_given_wk3,
    Chargers_wk2_vs_wk3$redzone_conversion_rate_allowed_wk3,
    Chargers_wk2_vs_wk3$sacks_forced_wk3,
    Chargers_wk2_vs_wk3$sack_forced_rate_wk3,
    Chargers_wk2_vs_wk3$pressures_wk3,
    Chargers_wk2_vs_wk3$pressure_rate_wk3,
    Chargers_wk2_vs_wk3$incomplete_pass_forced_rate_wk3,
    Chargers_wk2_vs_wk3$third_down_conversion_rate_allowed_wk3,
    Chargers_wk2_vs_wk3$EPA_given_up_per_play_wk3,
    Chargers_wk2_vs_wk3$EPA_passing_given_per_play_wk3,
    Chargers_wk2_vs_wk3$EPA_rushing_given_per_play_wk3,
    Chargers_wk2_vs_wk3$QB_EPA_given_per_play_wk3,
    Chargers_wk2_vs_wk3$def_success_rate_wk3,
    Chargers_wk2_vs_wk3$def_rush_success_rate_wk3,
    Chargers_wk2_vs_wk3$def_pass_success_rate_wk3,
    Chargers_wk2_vs_wk3$def_average_air_yards_depth_wk3),
  
  Difference = c(
    # Offense
    Chargers_wk2_vs_wk3$points_change_from_week_2,
    Chargers_wk2_vs_wk3$yards_change_from_week_2,
    Chargers_wk2_vs_wk3$passing_yards_change_from_week_2,
    Chargers_wk2_vs_wk3$rushing_yards_change_from_week_2,
    Chargers_wk2_vs_wk3$offensive_turnovers_change_from_week_2,
    Chargers_wk2_vs_wk3$turnover_differential_change_from_week_2,
    Chargers_wk2_vs_wk3$Net_EPA_change_from_week_2,
    Chargers_wk2_vs_wk3$Net_EPA_per_play_change_from_week_2,
    Chargers_wk2_vs_wk3$EPA_generated_change_from_week_2,
    Chargers_wk2_vs_wk3$EPA_passing_change_from_week_2,
    Chargers_wk2_vs_wk3$EPA_rushing_change_from_week_2,
    Chargers_wk2_vs_wk3$QB_EPA_change_from_week_2,
    Chargers_wk2_vs_wk3$redzone_EPA_change_from_week_2,
    Chargers_wk2_vs_wk3$redzone_conversion_rate_change_from_week_2,
    Chargers_wk2_vs_wk3$sacks_allowed_change_from_week_2,
    Chargers_wk2_vs_wk3$sack_rate_change_from_week_2,
    Chargers_wk2_vs_wk3$pressures_allowed_change_from_week_2,
    Chargers_wk2_vs_wk3$pressure_allowed_rate_change_from_week_2,
    Chargers_wk2_vs_wk3$incomplete_pass_rate_change_from_week_2,
    Chargers_wk2_vs_wk3$third_down_conversion_rate_change_from_week_2,
    Chargers_wk2_vs_wk3$EPA_generated_per_play_change_from_week_2,
    Chargers_wk2_vs_wk3$EPA_passing_per_play_change_from_week_2,
    Chargers_wk2_vs_wk3$EPA_rushing_per_play_change_from_week_2,
    Chargers_wk2_vs_wk3$QB_EPA_per_play_change_from_week_2,
    Chargers_wk2_vs_wk3$off_success_rate_change_from_week_2,
    Chargers_wk2_vs_wk3$off_rush_success_rate_change_from_week_2,
    Chargers_wk2_vs_wk3$off_pass_success_rate_change_from_week_2,
    Chargers_wk2_vs_wk3$off_average_air_yards_depth_change_from_week_2,
    
    # Defense
    Chargers_wk2_vs_wk3$defensive_plays_change_from_week_2,
    Chargers_wk2_vs_wk3$yards_allowed_change_from_week_2,
    Chargers_wk2_vs_wk3$passing_yards_allowed_change_from_week_2,
    Chargers_wk2_vs_wk3$rushing_yards_allowed_change_from_week_2,
    Chargers_wk2_vs_wk3$defensive_takeaways_change_from_week_2,
    Chargers_wk2_vs_wk3$EPA_given_up_change_from_week_2,
    Chargers_wk2_vs_wk3$EPA_passing_given_change_from_week_2,
    Chargers_wk2_vs_wk3$EPA_rushing_given_change_from_week_2,
    Chargers_wk2_vs_wk3$QB_EPA_given_change_from_week_2,
    Chargers_wk2_vs_wk3$redzone_EPA_given_change_from_week_2,
    Chargers_wk2_vs_wk3$redzone_conversion_rate_allowed_change_from_week_2,
    Chargers_wk2_vs_wk3$sacks_forced_change_from_week_2,
    Chargers_wk2_vs_wk3$sack_forced_rate_change_from_week_2,
    Chargers_wk2_vs_wk3$pressures_change_from_week_2,
    Chargers_wk2_vs_wk3$pressure_rate_change_from_week_2,
    Chargers_wk2_vs_wk3$incomplete_pass_forced_rate_change_from_week_2,
    Chargers_wk2_vs_wk3$third_down_conversion_rate_allowed_change_from_wk2,
    Chargers_wk2_vs_wk3$EPA_given_up_per_play_change_from_week_2,
    Chargers_wk2_vs_wk3$EPA_passing_given_per_play_change_from_week_2,
    Chargers_wk2_vs_wk3$EPA_rushing_given_per_play_change_from_week_2,
    Chargers_wk2_vs_wk3$QB_EPA_given_per_play_change_from_week_2,
    Chargers_wk2_vs_wk3$def_success_rate_change_from_week_2,
    Chargers_wk2_vs_wk3$def_rush_success_rate_change_from_week_2,
    Chargers_wk2_vs_wk3$def_pass_success_rate_change_from_week_2,
    Chargers_wk2_vs_wk3$def_average_air_yards_depth_change_from_week_2))%>%
  mutate(
    Direction = case_when(
      Difference > 0 ~ "↑",
      Difference < 0 ~ "↓",
      TRUE ~ "→"))
```

Heatmap
```{r}
Chargers_wk2_wk3_gt = Chargers_wk2_wk3_table%>%
  gt()%>%
  tab_header(
    title = "Chargers: Week 2 vs. Week 3",
    subtitle = "Positive differences indicate improved performance")%>%
  cols_label(
    Stat = "Statistic",
    week_2 = "Week 2",
    Week_3 = "Week 3",
    Difference = "Difference",
    Direction = "Direction")%>%
  fmt_number(
    columns = c(week_2, Week_3, Difference),
    decimals = 2)%>%
  cols_align(
    align = "center",
    columns = c(week_2, Week_3, Difference, Direction))%>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(everything()))%>%
  tab_style(
    style = cell_text(size = px(20), weight = "bold"),
    locations = cells_body(columns = Direction))%>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = Stat))%>%
  opt_row_striping()%>%
  opt_table_font(
    font = list(
      google_font("Roboto"),
      default_fonts()))%>%
  tab_options(
    table.font.size = px(14),
    heading.title.font.size = px(22),
    heading.subtitle.font.size = px(14))

Chargers_wk2_wk3_gt 
```

Saving table
```{r}
gtsave(
  Chargers_wk2_wk3_gt,
  "C:/Personal Projects/Statistically Speaking/Chargers_wk2_wk3.png")
```

Comparing our performance over 3 weeks to the league's average 

```{r}
league_wk_1 = league%>%
  filter(week == 1)

team_points_wk_1 = league_wk_1%>%
  group_by(game_id)%>%
  summarise(
    home_team = first(home_team),
    away_team = first(away_team),
    home_points = max(total_home_score, na.rm = TRUE),
    away_points = max(total_away_score, na.rm = TRUE),
    .groups = "drop")

team_points_wk_1 = bind_rows(
  team_points_wk_1 %>%
    transmute(
      team = home_team,
      points = home_points),
  
  team_points_wk_1 %>%
    transmute(
      team = away_team,
      points = away_points))
#Offense
offense_wk_1 = league_wk_1%>%
  filter(
    !is.na(posteam),
    !play_type %in% c(
      "kickoff",
      "extra_point",
      "field_goal"))%>%
  group_by(posteam)%>%
  summarise(
    
    yards = sum(yards_gained, na.rm = TRUE),
    
    passing_yards = sum(passing_yards, na.rm = TRUE),
    
    rushing_yards = sum(rushing_yards, na.rm = TRUE),
    
    offensive_turnovers = sum(
      interception + fumble_lost,
      na.rm = TRUE),
    
    EPA_generated = sum(epa, na.rm = TRUE),
    
    EPA_passing = sum(
      epa[play_type == "pass"],
      na.rm = TRUE),
    
    EPA_rushing = sum(
      epa[play_type == "run"],
      na.rm = TRUE),
    
    QB_EPA = sum(qb_epa, na.rm = TRUE),
    
    redzone_EPA = sum(
      epa[redzone == 1],
      na.rm = TRUE),
    
    redzone_drives = n_distinct(
      drive[redzone == 1]),
    
    redzone_touchdowns = n_distinct(
      drive[redzone == 1 & touchdown == 1]),
    
    redzone_conversion_rate =
      redzone_touchdowns / redzone_drives,
    
    sacks_allowed = sum(sack, na.rm = TRUE),
    
    pressures_allowed = sum(qb_hit, na.rm = TRUE),
    
    incomplete_passes = sum(incomplete_pass, na.rm = TRUE),
    
    offensive_plays = n(),
    
    pass_plays = sum(
      play_type == "pass",
      na.rm = TRUE),
    
    rush_plays = sum(
      play_type == "run",
      na.rm = TRUE),
    
    sack_rate =
      sacks_allowed / pass_plays,
    
    pressure_allowed_rate = pressures_allowed / pass_plays, 
    
    incomplete_pass_rate =
      incomplete_passes / pass_plays,
    
    third_downs_faced =
      sum(down == 3, na.rm = TRUE),
    
    third_down_conversion_rate =
      sum(third_down_converted, na.rm = TRUE) /
      third_downs_faced,
    
    EPA_generated_per_play =
      EPA_generated / offensive_plays,
    
    EPA_passing_per_play =
      EPA_passing / pass_plays,
    
    EPA_rushing_per_play =
      EPA_rushing / rush_plays,
    
    QB_EPA_per_play =
      QB_EPA / pass_plays,
    
    off_success_rate =
      sum(success, na.rm = TRUE) /
      offensive_plays,
    
    off_rush_success_rate =
      sum(success[play_type == "run"], na.rm = TRUE) /
      rush_plays,
    
    off_pass_success_rate =
      sum(success[play_type == "pass"], na.rm = TRUE) /
      pass_plays,
    
    off_average_air_yards_depth =
      mean(air_yards, na.rm = TRUE),
    
    .groups = "drop")%>%
  rename(team = posteam)

#Defense
defense_wk_1 = league_wk_1%>%
  filter(
    !is.na(defteam),
    !play_type %in% c(
      "kickoff",
      "extra_point",
      "field_goal"))%>%
  group_by(defteam)%>%
  summarise(
    
    yards_allowed = sum(
      yards_gained,
      na.rm = TRUE),
    
    passing_yards_allowed = sum(
      passing_yards,
      na.rm = TRUE),
    
    rushing_yards_allowed = sum(
      rushing_yards,
      na.rm = TRUE),
    
    defensive_takeaways = sum(
      interception + fumble_forced,
      na.rm = TRUE),
    
    EPA_given_up = sum(
      epa,
      na.rm = TRUE),
    
    EPA_passing_given = sum(
      epa[play_type == "pass"],
      na.rm = TRUE),
    
    EPA_rushing_given = sum(
      epa[play_type == "run"],
      na.rm = TRUE),
    
    QB_EPA_given = sum(
      qb_epa,
      na.rm = TRUE),
    
    redzone_EPA_given = sum(
      epa[redzone == 1],
      na.rm = TRUE),
    
    redzone_drives_allowed = n_distinct(
      drive[redzone == 1]),
    
    redzone_touchdowns_allowed = n_distinct(
      drive[redzone == 1 & touchdown == 1]),
    
    redzone_conversion_rate_allowed =
      redzone_touchdowns_allowed /
      redzone_drives_allowed,
    
    sacks_forced = sum(
      sack,
      na.rm = TRUE),
    
    pressures = sum(qb_hit, na.rm = TRUE),
    
    incomplete_passes_forced = sum(
      incomplete_pass,
      na.rm = TRUE),
    
    defensive_plays = n(),
    
    pass_plays_faced = sum(
      play_type == "pass",
      na.rm = TRUE),
    
    rush_plays_faced = sum(
      play_type == "run",
      na.rm = TRUE),
    
    sack_forced_rate =
      sacks_forced / pass_plays_faced,
    
    pressure_rate = pressures / pass_plays_faced,
    
    incomplete_pass_forced_rate =
      incomplete_passes_forced /
      pass_plays_faced,
    
    third_downs_forced =
      sum(down == 3, na.rm = TRUE),
    
    third_down_conversion_rate_allowed =
      sum(third_down_converted, na.rm = TRUE) /
      third_downs_forced,
    
    EPA_given_up_per_play =
      EPA_given_up / defensive_plays,
    
    EPA_passing_given_per_play =
      EPA_passing_given / pass_plays_faced,
    
    EPA_rushing_given_per_play =
      EPA_rushing_given / rush_plays_faced,
    
    QB_EPA_given_per_play =
      QB_EPA_given / pass_plays_faced,
    
    def_success_rate =
      1 - (
        sum(success, na.rm = TRUE) /
        defensive_plays),
    
    def_rush_success_rate =
      1 - (
        sum(success[play_type == "run"], na.rm = TRUE) /
        rush_plays_faced),
    
    def_pass_success_rate =
      1 - (
        sum(success[play_type == "pass"], na.rm = TRUE) /
        pass_plays_faced),
    
    def_average_air_yards_depth =
      mean(air_yards, na.rm = TRUE),
    
    .groups = "drop")%>%
  rename(team = defteam)


```

Joined Data
```{r}
league_wk_1_summary = offense_wk_1%>%
  left_join(
    defense_wk_1,
    by = "team")%>%
  left_join(
    team_points_wk_1,
    by = "team")%>%
  mutate(
    Net_EPA =
      EPA_generated - EPA_given_up,
    
    Net_EPA_per_play =
      EPA_generated_per_play -
      EPA_given_up_per_play,
    
    turnover_differential =
      defensive_takeaways - offensive_turnovers)%>%
  mutate(
    points_rank = min_rank(desc(points)),
    
    offensive_plays_rank = min_rank(desc(offensive_plays)),
    
    defensive_plays_rank = min_rank((defensive_plays)),
    
    yards_rank = min_rank(desc(yards)),
    
    passing_yards_rank = min_rank(desc(passing_yards)),
    
    rushing_yards_rank = min_rank(desc(rushing_yards)),
    
    offensive_turnovers_rank = min_rank(offensive_turnovers),
    
    EPA_generated_rank = min_rank(desc(EPA_generated)),
    
    EPA_passing_rank = min_rank(desc(EPA_passing)),
    
    EPA_rushing_rank = min_rank(desc(EPA_rushing)),
    
    QB_EPA_rank = min_rank(desc(QB_EPA)),
    
    redzone_EPA_rank = min_rank(desc(redzone_EPA)),
    
    redzone_conversion_rate_rank = min_rank(desc(redzone_conversion_rate)),
    
    sacks_allowed_rank = min_rank(sacks_allowed),
    
    sack_rate_rank = min_rank(sack_rate),
    
    pressures_allowed_rank = min_rank(pressures_allowed),
    
    pressure_allowed_rate_rank = min_rank(pressure_allowed_rate),
    
    incomplete_pass_rate_rank = min_rank(incomplete_pass_rate),
    
    third_down_conversion_rate_rank = min_rank(desc(third_down_conversion_rate)),
    
    EPA_generated_per_play_rank = min_rank(desc(EPA_generated_per_play)),
    
    EPA_passing_per_play_rank = min_rank(desc(EPA_passing_per_play)),
    
    EPA_rushing_per_play_rank = min_rank(desc(EPA_rushing_per_play)),
    
    QB_EPA_per_play_rank = min_rank(desc(QB_EPA_per_play)),
    
    off_success_rate_rank = min_rank(desc(off_success_rate)),
    
    off_rush_success_rate_rank = min_rank(desc(off_rush_success_rate)),
    
    off_pass_success_rate_rank = min_rank(desc(off_pass_success_rate)),
    
    off_average_air_yards_depth_rank = min_rank(desc(off_average_air_yards_depth)),
    
    yards_allowed_rank = min_rank(yards_allowed),
    
    passing_yards_allowed_rank = min_rank((passing_yards_allowed)),
    
    rushing_yards_allowed_rank = min_rank((rushing_yards_allowed)),
    
    defensive_takeaways_rank = min_rank(desc(defensive_takeaways)),
    
    EPA_given_up_rank = min_rank(EPA_given_up),
    
    EPA_passing_given_rank = min_rank(EPA_passing_given),
    
    EPA_rushing_given_rank = min_rank(EPA_rushing_given),
    
    QB_EPA_given_rank = min_rank(QB_EPA_given),
    
    redzone_EPA_given_rank = min_rank(redzone_EPA_given),
    
    redzone_conversion_rate_allowed_rank = min_rank(redzone_conversion_rate_allowed),
    
    sacks_forced_rank = min_rank(desc(sacks_forced)),
    
    sack_forced_rate_rank = min_rank(desc(sack_forced_rate)),
    
    pressures_rank = min_rank(desc(pressures)),
    
    pressure_rate_rank = min_rank(desc(pressure_rate)),
    
    incomplete_pass_forced_rate_rank = min_rank(desc(incomplete_pass_forced_rate)),
    
    third_down_conversion_rate_allowed_rank = min_rank(desc(third_down_conversion_rate_allowed)),
    
    EPA_given_up_per_play_rank = min_rank(EPA_given_up_per_play),
    
    EPA_passing_given_per_play_rank = min_rank(EPA_passing_given_per_play),
     
    EPA_rushing_given_per_play_rank = min_rank(EPA_rushing_given_per_play),
    
    QB_EPA_given_per_play_rank = min_rank(QB_EPA_given_per_play),
    
    def_success_rate_rank = min_rank(desc(def_success_rate)),
    
    def_rush_success_rate_rank = min_rank(desc(def_rush_success_rate)),
    
    def_pass_success_rate_rank = min_rank(desc(def_pass_success_rate)),
    
    def_average_air_yards_depth_rank = min_rank((def_average_air_yards_depth)),
    
    Net_EPA_rank = min_rank(desc(Net_EPA)),
    
    Net_EPA_per_play_rank = min_rank(desc(Net_EPA_per_play)),
    
    turnover_differential_rank = min_rank(desc(turnover_differential)))


```


Cleaning it up

```{r}
league_wk_1_summary = league_wk_1_summary%>%
  select(team, 
         points, 
         points_rank, 
         offensive_plays, 
         offensive_plays_rank, 
         yards,
         yards_rank, 
         passing_yards, 
         passing_yards_rank, 
         rushing_yards, 
         rushing_yards_rank, 
         turnover_differential, 
         turnover_differential_rank, 
         offensive_turnovers, 
         offensive_turnovers_rank, 
         defensive_plays, 
         defensive_takeaways, 
         defensive_takeaways_rank, 
         Net_EPA, 
         Net_EPA_rank, 
         Net_EPA_per_play, 
         Net_EPA_per_play_rank, 
         EPA_generated, 
         EPA_generated_rank, 
         EPA_passing, 
         EPA_passing_rank, 
         EPA_rushing, 
         EPA_rushing_rank, 
         QB_EPA, 
         QB_EPA_rank,
         redzone_EPA,
         redzone_EPA_rank,
         redzone_conversion_rate,
         redzone_conversion_rate_rank,
         sacks_allowed,
         sacks_allowed_rank,
         sack_rate,
         sack_rate_rank,
         pressures_allowed,
         pressures_allowed_rank,
         pressure_allowed_rate,
         pressure_allowed_rate_rank,
         incomplete_pass_rate,
         incomplete_pass_rate_rank,
         third_down_conversion_rate,
         third_down_conversion_rate_rank,
         EPA_passing_per_play, 
         EPA_passing_per_play_rank, 
         EPA_rushing_per_play, 
         EPA_rushing_per_play_rank, 
         QB_EPA_per_play, 
         QB_EPA_per_play_rank, 
         EPA_generated_per_play, 
         EPA_generated_per_play_rank,
         off_success_rate,
         off_success_rate_rank,
         off_rush_success_rate,
         off_rush_success_rate_rank,
         off_pass_success_rate,
         off_pass_success_rate_rank,
         off_average_air_yards_depth,
         off_average_air_yards_depth_rank,
         yards_allowed,
         yards_allowed_rank,
         passing_yards_allowed,
         passing_yards_allowed_rank,
         rushing_yards_allowed,
         rushing_yards_allowed_rank,
         EPA_given_up, 
         EPA_given_up_rank,
         EPA_passing_given,
         EPA_passing_given_rank,
         EPA_rushing_given,
         EPA_rushing_given_rank,
         QB_EPA_given,
         QB_EPA_given_rank,
         redzone_EPA_given,
         redzone_EPA_given_rank,
         redzone_conversion_rate_allowed,
         redzone_conversion_rate_allowed_rank,
         sacks_forced,
         sacks_forced_rank,
         sack_forced_rate,
         sack_forced_rate_rank,
         pressures,
         pressures_rank,
         pressure_rate,
         pressure_rate_rank, 
         incomplete_pass_forced_rate,
         incomplete_pass_forced_rate_rank,
         third_down_conversion_rate_allowed,
         third_down_conversion_rate_allowed_rank,
         EPA_given_up_per_play, 
         EPA_given_up_per_play_rank,
         EPA_passing_given_per_play,
         EPA_passing_given_per_play_rank,
         EPA_rushing_given_per_play,
         EPA_rushing_given_per_play_rank,
         QB_EPA_given_per_play,
         QB_EPA_given_per_play_rank,
         def_success_rate,
         def_success_rate_rank,
         def_rush_success_rate,
         def_rush_success_rate_rank,
         def_pass_success_rate,
         def_pass_success_rate_rank,
         def_average_air_yards_depth,
         def_average_air_yards_depth_rank)

```

Joining with Week 2 and 3 Summary

```{r}

all_stats = c(
  offensive_stats,
  defensive_stats)

league_through_wk3 = bind_rows(
  
  league_wk_1_summary%>%
    select(team, all_of(all_stats))%>%
    mutate(week = "Wk1"),
  
  league_wk_2_summary%>%
    select(team, all_of(all_stats))%>%
    mutate(week = "Wk2"),
  
  league_wk_3_summary%>%
    select(team, all_of(all_stats))%>%
    mutate(week = "Wk3"))

chargers_avg = league_through_wk3%>%
  filter(team == "LAC")%>%
  pivot_longer(
    cols = all_of(all_stats),
    names_to = "stat",
    values_to = "value")%>%
  select(week, stat, value)%>%
  pivot_wider(
    names_from = week,
    values_from = value,
    names_glue = "Chargers_{week}")%>%
  mutate(
    Chargers_Season_Avg =
      rowMeans(
        select(., Chargers_Wk1, Chargers_Wk2, Chargers_Wk3),
        na.rm = TRUE))


league_avg = league_through_wk3%>%
  group_by(week)%>%
  summarise(
    
    across(
      all_of(all_stats),
      ~ mean(.x, na.rm = TRUE)),
    
    .groups = "drop")%>%
  pivot_longer(
    cols = -week,
    names_to = "stat",
    values_to = "value")%>%
  pivot_wider(
    names_from = week,
    values_from = value,
    names_glue = "League_{week}")%>%
  mutate(
    League_Season_Avg =
      rowMeans(
        select(
          .,
          League_Wk1,
          League_Wk2,
          League_Wk3),
        na.rm = TRUE))


league_vs_chargers_season = league_avg%>%
  left_join(
    chargers_avg,
    by = "stat")%>%
  select(
    stat,
    League_Wk1,
    League_Wk2,
    League_Wk3,
    League_Season_Avg,
    Chargers_Wk1,
    Chargers_Wk2,
    Chargers_Wk3,
    Chargers_Season_Avg)

```









