---
title: "NFL Passing Analysis Cleaning"
author: "Josh Garzaniti"
date: "2026-05-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install and load in needed libraries
```{r}
library(nflfastR)
library(nflreadr)
library(tidyverse)
library(stringr)
library(ggplot2)
library(ggdark)
library(caret)
library(xgboost)
library(randomForest)
library(ggrepel)
library(nflplotR)
library(mice)
library(lme4)
library(here)
library(DT)
library(htmlwidgets)
```
Load in our play by play data
```{r}
pbp = load_pbp(2026)
```

First I want to separate into 2 dataframes: Chargers and the rest of the league

```{r}
chargers = pbp%>%
  filter(home_team == "LAC" | away_team == "LAC")

league = pbp

chargers_offense = chargers%>%
  filter(posteam == "LAC" & !play_type %in% c("kickoff", "extra_point", "field_goal"))

chargers_defense = chargers%>%
  filter(defteam == "LAC" & !play_type %in% c("kickoff", "extra_point", "field_goal"))

chargers_st = chargers%>%
  filter(play_type %in% c("kickoff", "extra_point", "field_goal"))
```


Week 1 vs Arizona Cardinals

```{r}
chargers_wk_1 = chargers%>%
  filter(away_team == "ARI")

chargers_offense_wk_1 = chargers%>%
  filter(posteam == "LAC" 
         & away_team == "ARI"
         & !play_type %in% c("kickoff", "extra_point", "field_goal"))

chargers_defense_wk_1 = chargers%>%
  filter(defteam == "LAC"
         & away_team == "ARI"
         & !play_type %in% c("kickoff", "extra_point", "field_goal"))

chargers_st_wk_1 = chargers%>%
  filter(play_type %in% c("kickoff", "extra_point", "field_goal")
         & away_team == "ARI")
```


Summary:

```{r}
total_wk1_points = max(chargers$total_home_score, na.rm = TRUE)

total_wk1_points
  
total_wk1_points_allowed = max(chargers$total_away_score, na.rm = TRUE)

total_wk1_points_allowed

total_wk1_yards = sum(chargers_offense_wk_1$yards_gained, na.rm = TRUE)

total_wk1_yards

total_wk1_yards_allowed = sum(chargers_defense_wk_1$yards_gained, na.rm = TRUE)

total_wk1_yards_allowed

total_wk1_turnover_dif = (sum(chargers_defense_wk_1$interception, chargers_defense_wk_1$fumble_forced, na.rm = TRUE) - sum(chargers_offense_wk_1$interception, chargers_offense_wk_1$fumble_lost, na.rm = TRUE))

total_wk1_turnover_dif

total_wk1_EPA_generated = sum(chargers_offense_wk_1$epa, na.rm = TRUE)

total_wk1_EPA_generated

total_wk1_EPA_given_up = sum(chargers_defense$epa, na.rm = TRUE)

total_wk1_EPA_given_up

total_wk1_Net_EPA = total_wk1_EPA_generated - total_wk1_EPA_given_up

total_wk1_Net_EPA

total_wk1_EPA_generated_per_play = total_wk1_EPA_generated/nrow(chargers_offense_wk_1)

total_wk1_EPA_generated_per_play

total_wk1_EPA_given_up_per_play = total_wk1_EPA_given_up/nrow(chargers_defense_wk_1)

total_wk1_EPA_given_up_per_play

total_wk1_Net_EPA_per_play = total_wk1_EPA_generated_per_play - total_wk1_EPA_given_up_per_play

total_wk1_Net_EPA_per_play

```

Chargers 14
Cardinals 26
Chargers yards: 268
Cardinals yards: 393
Turnover differential: -2
EPA generated: -15.8411
EPA allowed:4.18542
Net EPA: -20.02652
EPA generated per play: -0.2555016
EPA allowed per play:0.04982643
Net EPA per play: -0.305328


## How did the league do?

```{r}
league_wk_1 = league%>%
  filter(week == 1)
```

Generate Stats for each team:

Scoring Week 1
```{r}
team_points_wk_1 = league_wk_1%>%
  group_by(game_id)%>%
  summarise(
    home_team = first(home_team),
    away_team = first(away_team),
    home_points = max(total_home_score, na.rm = TRUE),
    away_points = max(total_away_score, na.rm = TRUE),
    .groups = "drop")

team_points_wk_1 = bind_rows(
  team_points_wk_1 %>%
    transmute(
      team = home_team,
      points = home_points),
  
  team_points_wk_1 %>%
    transmute(
      team = away_team,
      points = away_points))
```

Offensive Stats
```{r}
offense_wk_1 = league_wk_1%>%
  filter(
    !is.na(posteam),
    !play_type %in% c(
      "kickoff",
      "extra_point",
      "field_goal"))%>%
  group_by(posteam)%>%
  summarise(
    yards = sum(yards_gained, na.rm = TRUE),
    
    passing_yards = sum(passing_yards, na.rm = TRUE),
    
    rushing_yards = sum(rushing_yards, na.rm = TRUE),
    
    offensive_turnovers = sum(
      interception + fumble_lost,
      na.rm = TRUE),
    

    EPA_generated = sum(epa, na.rm = TRUE),
    

    EPA_passing = sum(
      epa[play_type == "pass"],
      na.rm = TRUE),
    
 
    EPA_rushing = sum(
      epa[play_type == "run"],
      na.rm = TRUE),
    
 
    QB_EPA = sum(qb_epa, na.rm = TRUE),
    
 
    offensive_plays = n(),
    
    pass_plays = sum(play_type == "pass", na.rm = TRUE),
    
    rush_plays = sum(play_type == "run", na.rm = TRUE),
    
 
    EPA_generated_per_play =
      EPA_generated / offensive_plays,
    
    EPA_passing_per_play =
      EPA_passing / pass_plays,
    
    EPA_rushing_per_play =
      EPA_rushing / rush_plays,
    
    QB_EPA_per_play =
      QB_EPA / pass_plays,
    
    off_success_rate = sum(success, na.rm = TRUE) / offensive_plays,
    
    off_rush_success_rate = sum(success[play_type == "run"], na.rm = TRUE) / rush_plays,
    
    off_pass_success_rate = sum(success[play_type == "pass"], na.rm = TRUE) / pass_plays,
    
    off_average_air_yards_depth = mean(air_yards, na.rm = TRUE),

    .groups = "drop")%>%
  rename(team = posteam)
```

Defensive Stats
```{r}
defense_wk_1 = league_wk_1%>%
  filter(
    !is.na(defteam),
    !play_type %in% c(
      "kickoff",
      "extra_point",
      "field_goal"))%>%
  group_by(defteam)%>%
  summarise(
    yards_allowed = sum(yards_gained, na.rm = TRUE),
    
    passing_yards_allowed = sum(passing_yards, na.rm = TRUE),
    
    rushing_yards_allowed = sum(rushing_yards, na.rm = TRUE),
    
    defensive_turnovers = sum(
  interception + fumble_forced,
  na.rm = TRUE),
  
  EPA_given_up= sum(epa, na.rm = TRUE),
    
  EPA_passing_given = sum(
      epa[play_type == "pass"],
      na.rm = TRUE),
    
 
    EPA_rushing_given = sum(
      epa[play_type == "run"],
      na.rm = TRUE),
    
 
    QB_EPA_given = sum(qb_epa, na.rm = TRUE),
    
 
    defensive_plays = n(),
    
    pass_plays_faced = sum(play_type == "pass", na.rm = TRUE),
    
    rush_plays_faced = sum(play_type == "run", na.rm = TRUE),
    
 
    EPA_given_up_per_play =
      EPA_given_up / defensive_plays,
    
    EPA_passing_given_per_play =
      EPA_passing_given / pass_plays_faced,
    
    EPA_rushing_given_per_play =
      EPA_rushing_given / rush_plays_faced,
    
    QB_EPA_given_per_play =
      QB_EPA_given / pass_plays_faced,
  
    def_success_rate =  1 - (sum(success, na.rm = TRUE) / defensive_plays),
    
    def_rush_success_rate = 1- (sum(success[play_type == "run"], na.rm = TRUE) / rush_plays_faced),
    
    def_pass_success_rate = 1- (sum(success[play_type == "pass"], na.rm = TRUE) / pass_plays_faced),
    
    def_average_air_yards_depth = mean(air_yards, na.rm = TRUE),
    
    .groups = "drop")%>%
  rename(team = defteam)
```

Joined Data
```{r}
league_wk_1_summary = offense_wk_1%>%
  left_join(
    defense_wk_1,
    by = "team")%>%
  left_join(
    team_points_wk_1,
    by = "team")%>%
  mutate(
    Net_EPA =
      EPA_generated - EPA_given_up,
    
    Net_EPA_per_play =
      EPA_generated_per_play -
      EPA_given_up_per_play,
    
    turnover_differential =
      defensive_turnovers - offensive_turnovers)%>%
  mutate(
    points_rank = min_rank(desc(points)),
    
    offensive_plays_rank = min_rank(desc(offensive_plays)),
    
    defensive_plays_rank = min_rank((defensive_plays)),
    
    yards_rank = min_rank(desc(yards)),
    
    passing_yards_rank = min_rank(desc(passing_yards)),
    
    rushing_yards_rank = min_rank(desc(rushing_yards)),
    
    offensive_turnovers_rank = min_rank(offensive_turnovers),
    
    EPA_generated_rank = min_rank(desc(EPA_generated)),
    
    EPA_passing_rank = min_rank(desc(EPA_passing)),
    
    EPA_rushing_rank = min_rank(desc(EPA_rushing)),
    
    QB_EPA_rank = min_rank(desc(QB_EPA)),
    
    EPA_generated_per_play_rank = min_rank(desc(EPA_generated_per_play)),
    
    EPA_passing_per_play_rank = min_rank(desc(EPA_passing_per_play)),
    
    EPA_rushing_per_play_rank = min_rank(desc(EPA_rushing_per_play)),
    
    QB_EPA_per_play_rank = min_rank(desc(QB_EPA_per_play)),
    
    off_success_rate_rank = min_rank(desc(off_success_rate)),
    
    off_rush_success_rate_rank = min_rank(desc(off_rush_success_rate)),
    
    off_pass_success_rate_rank = min_rank(desc(off_pass_success_rate)),
    
    off_average_air_yards_depth_rank = min_rank(desc(off_average_air_yards_depth)),
    
    yards_allowed_rank = min_rank(yards_allowed),
    
    passing_yards_allowed_rank = min_rank((passing_yards_allowed)),
    
    rushing_yards_allowed_rank = min_rank((rushing_yards_allowed)),
    
    defensive_turnovers_rank = min_rank(desc(defensive_turnovers)),
    
    EPA_given_up_rank = min_rank(EPA_given_up),
    
    EPA_passing_given_rank = min_rank(EPA_passing_given),
    
    EPA_rushing_given_rank = min_rank(EPA_rushing_given),
    
    QB_EPA_given_rank = min_rank(QB_EPA_given),
    
    EPA_given_up_per_play_rank = min_rank(EPA_given_up_per_play),
    
    EPA_passing_given_per_play_rank = min_rank(EPA_passing_given_per_play),
     
    EPA_rushing_given_per_play_rank = min_rank(EPA_rushing_given_per_play),
    
    QB_EPA_given_per_play_rank = min_rank(QB_EPA_given_per_play),
    
    def_success_rate_rank = min_rank(desc(def_success_rate)),
    
    def_rush_success_rate_rank = min_rank(desc(def_rush_success_rate)),
    
    def_pass_success_rate_rank = min_rank(desc(def_pass_success_rate)),
    
    def_average_air_yards_depth_rank = min_rank((def_average_air_yards_depth)),
    
    Net_EPA_rank = min_rank(desc(Net_EPA)),
    
    Net_EPA_per_play_rank = min_rank(desc(Net_EPA_per_play)),
    
    turnover_differential_rank = min_rank(desc(turnover_differential)))


```


Cleaning it up

```{r}
league_wk_1_summary = league_wk_1_summary%>%
  select(team, 
         points, 
         points_rank, 
         offensive_plays, 
         offensive_plays_rank, 
         yards, yards_rank, 
         passing_yards, 
         passing_yards_rank, 
         rushing_yards, 
         rushing_yards_rank, 
         turnover_differential, 
         turnover_differential_rank, 
         offensive_turnovers, 
         offensive_turnovers_rank, 
         defensive_plays, 
         defensive_turnovers, 
         defensive_turnovers_rank, 
         Net_EPA, 
         Net_EPA_rank, 
         Net_EPA_per_play, 
         Net_EPA_per_play_rank, 
         EPA_generated, 
         EPA_generated_rank, 
         EPA_passing, 
         EPA_passing_rank, 
         EPA_rushing, 
         EPA_rushing_rank, 
         QB_EPA, 
         QB_EPA_rank, 
         EPA_passing_per_play, 
         EPA_passing_per_play_rank, 
         EPA_rushing_per_play, 
         EPA_rushing_per_play_rank, 
         QB_EPA_per_play, 
         QB_EPA_per_play_rank, 
         EPA_generated_per_play, 
         EPA_generated_per_play_rank,
         off_success_rate,
         off_success_rate_rank,
         off_rush_success_rate,
         off_rush_success_rate_rank,
         off_pass_success_rate,
         off_pass_success_rate_rank,
         off_average_air_yards_depth,
         off_average_air_yards_depth_rank,
         EPA_given_up, 
         EPA_given_up_rank,
         EPA_passing_given,
         EPA_passing_given_rank,
         EPA_rushing_given,
         EPA_rushing_given_rank,
         QB_EPA_given,
         QB_EPA_given_rank,
         EPA_given_up_per_play, 
         EPA_given_up_per_play_rank,
         EPA_passing_given_per_play,
         EPA_passing_given_per_play_rank,
         EPA_rushing_given_per_play,
         EPA_rushing_given_per_play_rank,
         QB_EPA_given_per_play,
         QB_EPA_given_per_play_rank,
         def_success_rate,
         def_success_rate_rank,
         def_rush_success_rate,
         def_rush_success_rate_rank,
         def_pass_success_rate,
         def_pass_success_rate_rank,
         def_average_air_yards_depth,
         def_average_air_yards_depth_rank)
```


Chargers Data Table

```{r}
Chargers_Week_1_Summary = league_wk_1_summary %>%
  filter(team == "LAC")

DT::datatable(Chargers_Week_1_Summary)
```


Comparing Chargers Week 1 Adv Statistics to the League

```{r}
stats_to_compare = c(
  "points",
  "offensive_plays",
  "yards",
  "passing_yards",
  "rushing_yards",
  "turnover_differential",
  "offensive_turnovers",
  "defensive_plays",
  "defensive_turnovers",
  "Net_EPA",
  "Net_EPA_per_play",
  "EPA_generated",
  "EPA_passing",
  "EPA_rushing",
  "QB_EPA",
  "EPA_generated_per_play",
  "EPA_passing_per_play",
  "EPA_rushing_per_play",
  "QB_EPA_per_play",
  "off_success_rate",
  "off_rush_success_rate",
  "off_pass_success_rate",
  "off_average_air_yards_depth")

defensive_stats_to_compare = c(
  "EPA_given_up",
  "EPA_passing_given",
  "EPA_rushing_given",
  "QB_EPA_given",
  "EPA_given_up_per_play",
  "EPA_passing_given_per_play",
  "EPA_rushing_given_per_play",
  "QB_EPA_given_per_play",
  "def_success_rate",
  "def_rush_success_rate",
  "def_pass_success_rate",
  "def_average_air_yards_depth")

Chargers_vs_League = league_wk_1_summary%>%
  filter(team == "LAC")%>%
  mutate(
    across(
      all_of(stats_to_compare),
      ~ .x - mean(league_wk_1_summary[[cur_column()]], na.rm = TRUE),
      .names = "{.col}_vs_league_avg"),
    
    across(
      all_of(defensive_stats_to_compare),
      ~ mean(league_wk_1_summary[[cur_column()]], na.rm = TRUE) - .x,
      .names = "{.col}_vs_league_avg"))
```

Data Table of Chargers week 1 stats
```{r}
DT::datatable(
  Chargers_vs_League,
  options = list(
    scrollX = TRUE,
    pageLength = 10))
```

Chargers vs League Stats Heatmap week 1:

Offense
```{r}
offense_stats = c(
  "points",
  "yards",
  "passing_yards",
  "rushing_yards",
  "turnover_differential",
  "Net_EPA",
  "Net_EPA_per_play",
  "EPA_passing",
  "EPA_rushing",
  "QB_EPA",
  "EPA_passing_per_play",
  "EPA_rushing_per_play",
  "QB_EPA_per_play",
  "off_success_rate",
  "off_rush_success_rate",
  "off_pass_success_rate",
  "off_average_air_yards_depth")

heatmap_data = league_wk_1_summary%>%
  select(team, all_of(offense_stats))%>%
  pivot_longer(
    cols = -team,
    names_to = "stat",
    values_to = "value")%>%
  group_by(stat)%>%
  mutate(
    league_avg = mean(value, na.rm = TRUE),
    vs_league_avg = value - league_avg)%>%
  ungroup()
```

Defense
```{r}
defensive_stats = c(
  "EPA_passing_given",
  "EPA_rushing_given",
  "QB_EPA_given",
  "EPA_passing_given_per_play",
  "EPA_rushing_given_per_play",
  "QB_EPA_given_per_play",
  "def_success_rate",
  "def_rush_success_rate",
  "def_pass_success_rate",
  "def_average_air_yards_depth")

defensive_lower_is_better = c(
  "EPA_passing_given",
  "EPA_rushing_given",
  "QB_EPA_given",
  "EPA_passing_given_per_play",
  "EPA_rushing_given_per_play",
  "QB_EPA_given_per_play",
  "def_average_air_yards_depth")

heatmap_stats = c(
  offense_stats,
  defensive_stats)

heatmap_data = league_wk_1_summary%>%
  select(team, all_of(heatmap_stats))%>%
  pivot_longer(
    cols = -team,
    names_to = "stat",
    values_to = "value")%>%
  group_by(stat) %>%
  mutate(
    league_avg = mean(value, na.rm = TRUE),
    vs_league_avg = value - league_avg)%>%
  ungroup() %>%
  mutate(
    performance_vs_avg = case_when(
  stat %in% defensive_lower_is_better ~ -vs_league_avg,
  TRUE ~ vs_league_avg))%>%
  group_by(stat)%>%
  mutate(
    z_score = (
      performance_vs_avg -
        mean(performance_vs_avg, na.rm = TRUE)) /
      sd(performance_vs_avg, na.rm = TRUE))%>%
  ungroup()
```

Offensive Heatmap
```{r}
offense_heatmap_data = heatmap_data%>%
  filter(stat %in% offense_stats)

ggplot(
  offense_heatmap_data,
  aes(
    x = team,
    y = stat,
    fill = z_score))+
  
  geom_tile(color = "grey20")+
  
  scale_fill_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = 0)+
  
  labs(
    title = "NFL Week 1 Offensive Performance vs. League Average",
    subtitle = "Positive values indicate above-average performance",
    x = "Team",
    y = "Statistic",
    fill = "Vs. League Avg.")+
  
  theme_minimal()+
  theme(
    plot.background = element_rect(fill = "black", color = NA),
    panel.background = element_rect(fill = "black", color = NA),
    panel.grid = element_blank(),
    
    text = element_text(color = "white"),
    axis.text = element_text(color = "white"),
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      color = "white"),
    axis.title = element_text(color = "white"),
    
    plot.title = element_text(
      color = "white",
      face = "bold"),
    plot.subtitle = element_text(color = "grey80"),
    
    legend.background = element_rect(fill = "black", color = NA),
    legend.text = element_text(color = "white"),
    legend.title = element_text(color = "white"))
```

Defensive heatmap

```{r}

defense_heatmap_data = heatmap_data%>%
  filter(stat %in% defensive_stats)

ggplot(
  defense_heatmap_data,
  aes(
    x = team,
    y = stat,
    fill = z_score))+
  
  geom_tile(color = "grey20")+
  
  scale_fill_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = 0)+
  
  labs(
    title = "NFL Week 1 Defensive Performance vs. League Average",
    subtitle = "Positive values indicate above-average performance",
    x = "Team",
    y = "Statistic",
    fill = "Vs. League Avg.")+
  
  theme_minimal()+
  theme(
    plot.background = element_rect(fill = "black", color = NA),
    panel.background = element_rect(fill = "black", color = NA),
    panel.grid = element_blank(),
    
    text = element_text(color = "white"),
    axis.text = element_text(color = "white"),
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      color = "white"),
    axis.title = element_text(color = "white"),
    
    plot.title = element_text(
      color = "white",
      face = "bold"),
    plot.subtitle = element_text(color = "grey80"),
    
    legend.background = element_rect(fill = "black", color = NA),
    legend.text = element_text(color = "white"),
    legend.title = element_text(color = "white"))
```





